---
editor_options: 
  markdown: 
    wrap: 72
---

\@\@ -5,114 +5,153 \@\@ date: "`r Sys.Date()`" output: html_document ---

```{r setup, include = FALSE}
# Load all required packages
library(ggplot2)
library(dplyr)
library(forcats)
library(tidyr)
library(stringr)
library(scales)
library(readr)
library(janitor)
library(corrplot)

# Global knit options
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
set.seed(2025)

```

```{r, include=FALSE}
library(scales)
# setup: load libraries and options
library(tidyverse)
library(readr)
library(haven)
library(janitor)
library(scales)
library(stringr)
library(dplyr)
library(tidyr)


knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
set.seed(2025)
```

## Introduction

In recent years, artificial intelligence (AI) has shifted from a source
of uncertainty and risk to a common feature of everyday life. From
writing assistants and chat bots to algorithmic recommendations in
education, finance, and entertainment, AI has become deeply embedded in
social and professional contexts. Understanding how people perceive its
benefits and risks is crucial for policymakers and organizations
designing AI systems that are equitable and trustworthy. In recent
years, artificial intelligence (AI) has undergone a profound shift in
public perception. Once seen primarily as a source of uncertainty or
even risk, AI is now becoming something people feel increasingly
comfortable with—if not socially or professionally expected to use. From
writing assistants and chatbots to algorithmic recommendations in
education, finance, and entertainment, AI has become an embedded feature
of everyday life.

To ground our analysis, we use the Public Attitudes to AI (PAI) data set
produced by the Ada Lovelace Institute (UK). This nationally
representative survey of approximately 4,000 adults captures attitudes
toward AI across multiple domains—such as hiring, healthcare, and
policing. To ground our analysis, we use the Public Attitudes to AI
dataset produced by the Ada Lovelace Institute (UK). This nationally
representative survey of \~4,000 adults provides perceptions across
multiple AI application domains. Our aim here is a reproducible initial
exploration: inspect structure and missingness, perform light cleaning,
compute simple summaries, and visualise perceived benefits of AI in
hiring.

Data set and Preprocessing This project uses the preprocessed version of
the data set provided by the Ada Lovelace Institute. The data have
already been cleaned, anonymized, and standardized by the research team,
including: \## Load data

1.  recording of categorical variables (e.g., sex, age_group);
2.  conversion of “benefit” items into binary 0/1 indicators; and
3.  removal of personally identifying information and invalid responses.

Because of this prior work, we do not replicate the raw data cleaning
here. Instead, we focus on exploratory data analysis (EDA) to identify
overall trends and patterns in public attitudes toward AI.

Research Questions RQ1: Which AI benefits are most widely perceived by
respondents? RQ2: How do perceptions of AI benefits vary by gender and
age? RQ3: What patterns in the data suggest directions for further
analysis later in the semester?

## Loading raw data set

```{r}
ai_path <- "how-do-people-feel-about-ai-main/262400371_Public Attitudes to AI Raw Data NA (1).csv"
if (!file.exists(ai_path)) stop("Data file not found: ", ai_path)
ai_survey1 <- read_csv(ai_path, show_col_types = FALSE)
head(ai_survey1, 10)
```

## Understanding the Variables

The Data set includes demographic variables such as age, sex, and region
along with multiple binary columns (beginning with 'benb\_')
representing whether respondents selected each perceived benefit. Below
we summarize the key variables used in this analysis

```{r}
tibble::tribble(
~variable,       ~description,                                    ~type,
"age",           "Respondent age in years",                       "Numeric",
"sex",           "Self-identified gender (Male/Female/Other)",    "Categorical",
"benb_*",        "Binary indicators for perceived AI benefits",   "Integer (0/1)",
"weight",        "Survey weight (if applicable)",                 "Numeric"
) %>%
knitr::kable(caption = "Key variables in the preprocessed dataset")
```

```{r}
## functions to simplify data tidying

## recoding technology labels
recode_technologies <- function(Technology_specific) {
  recode(Technology_specific,
                  "WEL" = "Welfare eligibility",
                "JOB" =  "Job eligibility",
                "CAN" =  "Cancer risk",
                "LOAN" = "Loan repayment risk", 
                "MP" =  "Mobile phone unlock",
                 "BC" = "Border control",
                 "PS" = "Police surveillance",

                "AW"  = "Autonomous weapons", 
                "DC" =  "Driverless cars",
                "RCA" = "Robotic care assistant",
               "RVC" ="Robotic vacuum cleaner",
                                                                         
               "SMC"=  "Targeted consumer ads",
                "SMP" = "Targeted political ads",
                                                                         
                "VASS"=  "Smart speaker",
                "VAH"=  "Virtual healthcare assistant",
                                 
              "CCR"=   "Climate change simulation",
                "VR"=  "Education simulations")
  
}
```

```{r}
## list of demographic variables of interest
demog <- ai_survey1 %>%
  haven::zap_labels() %>%
  dplyr::select(
  Respondent_ID, PV25_Weight,
  RS_Ethnicity_Reduced,
  CS_Age_Band,
  RS_Sex,
  RS_EducationLevel,
  DR_Socio_Eco
  ) %>%
  mutate(RS_Ethnicity_Combined = 
           dplyr::recode(RS_Ethnicity_Reduced,
             "1" = "White",
             "2" = "Asian or Asian British",
             "3" = "Black British, Caribbean or African",
             "4" = "Any other ethnic group"
           )) %>%
  mutate(CS_Age_Band = 
           dplyr::recode(CS_Age_Band,
             "1" = "18-24 yrs",
             "2" = "25-34 yrs",
             "3" = "35-44 yrs",
             "4" = "45-54 yrs",
             "5" = "55-64 yrs",
             "6" = "65-74 yrs",
             "7" = "75+ yrs"
           )) %>%
  mutate(RS_Sex = 
           dplyr::recode(RS_Sex,
             "1" = "Male",
             "2" = "Female")) %>%
  mutate(RS_EducationLevel = 
           dplyr::recode(RS_EducationLevel,
             "1" = "Degree level qualification(s)",
             "2" = "Non-degree level qualifications",
             "3" = "No academic or vocational qualifications")) %>%
  mutate(DR_Socio_Eco = 
           dplyr::recode(DR_Socio_Eco,
             "1" = "SEC1, 2",
             "2" = "SEC3",
             "3" = "SEC4",
             "4" = "SEC5",
             "5" = "SEC6, 7",
             "6" = "SEC8",
             "7" = "Students")) %>%
  pivot_longer(names_to = "Q_Demo",
               values_to = "A_Demo",
               RS_Ethnicity_Combined:DR_Socio_Eco)

# list of levels 
demo_list <- c("White",
  "Asian or Asian British",
  "Black British, Caribbean or African", 
  "Any other ethnic group", 
  
  "18-24 yrs",
  "25-34 yrs",
  "35-44 yrs",
  "45-54 yrs",
  "55-64 yrs",
  "65-74 yrs",
  "75+ yrs", 
  
  "Male",
  "Female",
  "Identify in another way",
  
  "Degree level qualification(s)",
  "Non-degree level qualifications",
  "No academic or vocational qualifications",
  
  "SEC1, 2",
  "SEC3",
  "SEC4",
  "SEC5",
  "SEC6, 7",
  "SEC8",
  "Students", 
  
  "Total" )
```

## Exploratory Data Analysis Age Distribution We start by visualizing the
age distribution to understand the demographic range represented in the
survey

```{r}
if ("age" %in% names(ai_survey1)) {
ggplot(filter(ai, !is.na(age)), aes(age)) +
geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
labs(title = "Age Distribution of Respondents", x = "Age (years)", y = "Count")
}
```

## Inspect variables and missingness

Before diving deeper, we can check for missing values and variables
types

```{r, eval = FALSE}
library(purrr)
```

```{r inspect-variables, eval = FALSE}

var_df <- tibble(
  var = names(ai_survey1)
)

print(var_df, n = 250)
```

## Identify and Prepare Key Columns We standardize variable names, detect
age and gender columns and identify all AI benefit-related columns
(benb\_). \## Basic cleaning and derived columns

```{r}
ai <- read_csv(ai_path, show_col_types = FALSE)
# Standardise names and add row id
ai_clean <- ai_survey1 %>%
  clean_names() %>%
  mutate(row_id = row_number())

ai_clean <- ai %>%
clean_names() %>%
mutate(row_id = row_number())

# Detect age column

age_col <- names(ai_clean)[str_detect(names(ai_clean), regex("age", ignore_case = TRUE))][1]
# Detect age column (common names cs_age, csage, cs_age_yrs, age)
age_col <- names(ai_clean)[str_detect(names(ai_clean), regex("\\b(cs_age|csage|age|age_years|cs_age_yrs)\\b", ignore_case = TRUE))] %>% .[1]
if (!is.na(age_col)) {
ai_clean <- ai_clean %>%
mutate(age = as.integer(.data[[age_col]])) %>%
filter(is.na(age) | age >= 17)
  ai_clean <- ai_clean %>%
    mutate(age = as.integer(.data[[age_col]])) %>%
    filter(is.na(age) | age >= 17)  # remove implausible 16 if present
}

# Detect sex/gender column

sex_col <- names(ai_clean)[str_detect(names(ai_clean), regex("sex|gender", ignore_case = TRUE))][1]
sex_col <- names(ai_clean)[str_detect(names(ai_clean), regex("^rs_sex$|sex|gender", ignore_case = TRUE))] %>% .[1]
if (!is.na(sex_col)) {
ai_clean <- ai_clean %>%
mutate(
sex_raw = as.character(.data[[sex_col]]),
sex = case_when(
sex_raw %in% c("1", "Male", "male", "M") ~ "Male",
sex_raw %in% c("2", "Female", "female", "F") ~ "Female",
TRUE ~ NA_character_
)
)
  ai_clean <- ai_clean %>%
    mutate(
      sex_raw = as.character(.data[[sex_col]]),
      sex = case_when(
        sex_raw %in% c("1", 1, "Male", "male", "M") ~ "Male",
        sex_raw %in% c("2", 2, "Female", "female", "F") ~ "Female",
        TRUE ~ NA_character_
      )
    )
}

# Detect benefit-related columns

benb_cols <- names(ai_clean)[str_detect(names(ai_clean), regex("^benb|benefit", ignore_case = TRUE))]
# glimpse(select(ai_clean, any_of(c("age", "sex", benb_cols))))

```

## Show top 10 variables with most missing variables

```{r}
ai_missing <- tibble(
variable = names(ai_clean),
missing = map_int(ai_clean, ~ sum(is.na(.))),
missing_pct = round(map_dbl(ai_clean, ~ mean(is.na(.))) * 100, 2)
) %>%
arrange(desc(missing_pct)) %>%
slice_head(n = 10)

knitr::kable(ai_missing, caption = "Top 10 Variables with the Highest Missingness")
# Detect BENB-like columns and coerce to 0/1 integers where possible
benb_cols <- names(ai_clean)[str_detect(names(ai_clean), regex("^benb|benb_job|benefit|benefits|benb_", ignore_case = TRUE))]
if (length(benb_cols) == 0) {
  benb_cols <- names(ai_clean)[str_detect(names(ai_clean), regex("benefit|benb|job", ignore_case = TRUE))]
}

if (length(benb_cols) > 0) {
  ai_clean <- ai_clean %>%
    mutate(across(all_of(benb_cols), ~ {
      x <- .
      # character encodings -> 1/0
      if (is.character(x)) {
        xl <- str_to_lower(str_trim(x))
        as.integer(if_else(xl %in% c("yes","y","1","selected","true","t"), 1L,
                   if_else(xl %in% c("no","n","0","false","f"), 0L, NA_integer_)))
      } else if (is.logical(x)) {
        as.integer(x)
      } else {
        # numeric-like: keep 1/0, else NA
        xi <- suppressWarnings(as.integer(x))
        if_else(is.na(xi), NA_integer_, xi)
      }
    }))
}
# quick check
# glimpse(select(ai_clean, any_of(c("age","sex", benb_cols))))
```

## Derive Overall Benefit Tallies (Unweighted)

```{r}
# define which columns are your 0/1 benefit flags
benefit_cols <- grep("^benb_", names(ai_clean), value = TRUE)
# benefit_cols

df_long <- ai_clean %>%
  mutate(row_id = dplyr::row_number()) %>%
  tidyr::pivot_longer(
    cols = dplyr::all_of(benefit_cols),
    names_to   = "benefit_raw",
    values_to  = "selected",
    values_transform = list(selected = as.integer),  # coerce if needed
    values_drop_na   = TRUE
  ) %>%
  dplyr::filter(selected == 1L) %>%
  dplyr::mutate(benefit = benefit_raw) %>%
  dplyr::count(benefit) %>%
  dplyr::mutate(pct = n / sum(n))
```

```{r, eval= FALSE}
 # ai_clean %>% 
 #  dplyr::summarise(across(everything(), ~ all(. %in% c(0,1), na.rm = TRUE))) %>%
 # t() %>%
  # tibble::as_tibble(rownames = "var") %>%
  # dplyr::filter(value == TRUE)

```



```{r derive-overall-ben}
overall_ben <- ai_clean %>%
  tidyr::pivot_longer(all_of(benb_cols), names_to = "benefit", values_to = "selected",
                      values_transform = list(selected = as.integer),
                      values_drop_na = TRUE) %>%
  dplyr::filter(selected == 1L) %>%
  dplyr::count(benefit, sort = TRUE) %>%
  dplyr::mutate(pct = n / sum(n))
```

## Summary statistics

```{r summary-stats}
# Age summaries
if ("age" %in% names(ai_clean)) {
ai_clean %>%
summarise(
mean_age = mean(age, na.rm = TRUE),
median_age = median(age, na.rm = TRUE),
sd_age = sd(age, na.rm = TRUE),
min_age = min(age, na.rm = TRUE),
max_age = max(age, na.rm = TRUE)
)
  age_stats <- ai_clean %>%
    summarise(
      n = sum(!is.na(age)),
      mean_age = mean(age, na.rm = TRUE),
      median_age = median(age, na.rm = TRUE),
      sd_age = sd(age, na.rm = TRUE),
      min_age = min(age, na.rm = TRUE),
      max_age = max(age, na.rm = TRUE)
    )
  print(age_stats)
}

# Sex / gender distribution
if ("sex" %in% names(ai_clean)) {
ai_clean %>%
count(sex, sort = TRUE) %>%
mutate(percent = scales::percent(n / sum(n)))
  print(ai_clean %>% count(sex, sort = TRUE) %>% mutate(p = n / sum(n)))
}

# Top benefits preview
if (exists("overall_ben") && nrow(overall_ben) > 0) {
  print(overall_ben %>% arrange(desc(pct)) %>% slice_head(n = 10))
} ##### add explanation of the variable names
```

## Visualization

Age Histogram

```{r plots, fig.width=7, fig.height=4}
# Age histogram
if ("age" %in% names(ai_clean)) {
ggplot(filter(ai_clean, !is.na(age)), aes(age)) +
geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
labs(title = "Age Distribution", x = "Age", y = "Count")
  ggplot(filter(ai_clean, !is.na(age)), aes(age)) +
    geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
    labs(title = "Age distribution", x = "Age", y = "Count")
}

```

Most participants fall between their 20s and 50s, showing a good mix of
working-age adults. Fewer people over 70 took the survey, but the data
still includes a wide range of ages.

Gender Distribution

```{r}
# Gender bar chart

# Sex / gender bar
if ("sex" %in% names(ai_clean)) {
ggplot(filter(ai_clean, !is.na(sex)), aes(x = fct_infreq(sex))) +
geom_bar(fill = "darkorange") +
coord_flip() +
labs(title = "Gender Distribution", x = NULL, y = "Count")
  ggplot(filter(ai_clean, !is.na(sex)), aes(x = fct_infreq(sex))) +
    geom_bar(fill = "darkorange") +
    coord_flip() +
    labs(title = "Sex / Gender distribution", x = NULL, y = "Count")
}
```

The chart shows that survey responses were evenly split between male and
female participants, giving a balanced view of public attitudes toward
AI. This balance helps ensure that later comparisons by gender aren’t
biased by unequal sample sizes.

Top 8 Perceived Benefits of AI

```{r}
overall_ben <- overall_ben %>%
  mutate(
    benefit = str_to_lower(str_trim(benefit)),  # clean names first
    benefit_label = recode(benefit,
      "benb_mp01" = "Improved Job Matching",
      "benb_mp02" = "More Efficient Hiring",
      "benb_rvc03" = "Reduced Human Bias",
      "benb_can01" = "Career Advancement",
      "benb_ps01"  = "Personal Safety",
      "benb_vass02" = "Accessible Services",
      "benb_bc01"  = "Better Customer Service",
      "benb_rvc01" = "Reduced Cost",
      .default = benefit  # keeps unmapped names so nothing breaks
    )
  )

```

```{r}
# Benefits bar (top 8)
top8 <- overall_ben %>%
  arrange(desc(n)) %>%
  slice_head(n = 8) %>%
  mutate(
    benefit_clean = benefit %>%
      gsub("^benb_", "", .) %>%
      gsub("_", " ", .) %>%
      tools::toTitleCase()
  )

ggplot(top8, aes(x = n, y = reorder(benefit_clean, n))) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = n), hjust = -0.1, size = 3.7) +
  coord_cartesian(xlim = c(0, max(top8$n) * 1.15)) +
  labs(
    title = "Top 8 Benefits Selected",
    subtitle = "Raw number of respondents selecting each benefit",
    x = "Number of Respondents",
    y = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    axis.text.y = element_text(size = 10)
  )

## recode to show the actual benefit and not the coded version
```

### Edit this to be up to date with the actual benefits
The chart shows the top eight benefits people see in AI, with most
responses centered on work and efficiency. Improved Job Matching and
More Efficient Hiring lead the list, while others like Reduced Human
Bias and Career Advancement reflect optimism about fairness and
opportunity. Overall, respondents view AI as a tool that makes life
easier, safer, and a bit more convenient.

Correlation Heat Map

```{r}

library(ggcorrplot)

#select only the top 10 benefit columns
top10_ben <- overall_ben %>%
  arrange(desc(pct)) %>%
  slice_head(n = 10) %>%
  pull(benefit)

#correlation matrix
benb_matrix <- ai_clean %>%
  select(all_of(top10_ben)) %>%
  mutate(across(everything(), as.numeric)) %>%
  cor(use = "pairwise.complete.obs")

#replace any NAs with 0
benb_matrix[is.na(benb_matrix)] <- 0

#readable labels
benefit_labels <- c(
  "benb_mp01" = "Improved Job Matching",
  "benb_mp02" = "More Efficient Hiring",
  "benb_rvc03" = "Reduced Human Bias",
  "benb_can01" = "Career Advancement",
  "benb_ps01"  = "Personal Safety",
  "benb_vass02" = "Accessible Services",
  "benb_bc01"  = "Better Customer Service",
  "benb_rvc01" = "Reduced Cost",
  "benb_vr03"  = "Virtual Reality Support"
)

rownames(benb_matrix) <- benefit_labels[rownames(benb_matrix)]
colnames(benb_matrix) <- benefit_labels[colnames(benb_matrix)]

#plot larger and clearer with ggcorrplot
ggcorrplot(
  benb_matrix,
  type = "lower",
  lab = TRUE,         # show numeric values
  lab_size = 3.5,     # larger label size
  tl.cex = 10,        # enlarge text labels
  colors = c("steelblue", "brown", "yellow"),
  ggtheme = theme_minimal(base_size = 14),
  title = "Correlation Between Top 10 AI Benefit Selections"
) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16)
  )

```

This heatmap shows how the top ten perceived benefits of AI relate to
one another. The colors represent the strength of correlation between
benefits: lighter yellow tones indicate stronger positive relationships,
while darker brown areas show weaker or little association.

Overall, most benefits are positively correlated, meaning people who
view AI favorably in one area (like Improved Job Matching or More
Efficient Hiring) tend to also see advantages in related areas such as
Reduced Human Bias or Career Advancement. This suggests that respondents
often see AI’s impact as multi-dimensional—improving fairness,
opportunity, and convenience at the same time.

AI Benefits by Gender

```{r}
# --- Perceived AI Benefits by Gender (fixed version) ---
if (exists("benb_cols") && "sex" %in% names(ai_clean)) {

  gender_benefit <- ai_clean %>%
    select(sex, all_of(benb_cols)) %>%
    mutate(across(all_of(benb_cols), as.character)) %>%
    pivot_longer(-sex, names_to = "benefit", values_to = "selected") %>%
    mutate(
      selected = str_to_lower(str_trim(selected)),
      selected = case_when(
        selected %in% c("1","yes","y","true","selected") ~ 1L,
        selected %in% c("0","no","n","false","not selected") ~ 0L,
        TRUE ~ NA_integer_
      )
    ) %>%
    group_by(sex, benefit) %>%
    summarise(pct = mean(selected, na.rm = TRUE), .groups = "drop")

  # Top 8 by overall pct
  top8 <- overall_ben %>%
    arrange(desc(pct)) %>%
    slice_head(n = 8) %>%
    pull(benefit)

  # Pretty labels (add more as needed)
  benefit_labels <- c(
    "benb_mp01"  = "Improved Job Matching",
    "benb_mp02"  = "More Efficient Hiring",
    "benb_rvc03" = "Reduced Human Bias",
    "benb_can01" = "Career Advancement",
    "benb_ps01"  = "Personal Safety",
    "benb_vass02"= "Accessible Services",
    "benb_bc01"  = "Better Customer Service",
    "benb_rvc01" = "Reduced Cost"
  )

  plot_df <- gender_benefit %>%
    filter(benefit %in% top8) %>%
    mutate(
      benefit_label = recode(benefit, !!!benefit_labels, .default = benefit),
      # ensure ordering by overall pct (not within-sex)
      benefit_label = factor(benefit_label,
                             levels = recode(top8, !!!benefit_labels, .default = top8))
    )

  ggplot(plot_df, aes(x = benefit_label, y = pct, fill = sex)) +
    geom_col(position = "dodge") +
    coord_flip() +
    scale_y_continuous(labels = percent_format(accuracy = 1)) +
    scale_fill_manual(values = c("pink", "lightblue")) +
    labs(
      title = "Perceived benefits of AI in hiring (top items)",
      x = NULL,
      y = "Unweighted % selecting",
      fill = "Gender"
    ) +
    theme_minimal(base_size = 13) +
    theme(plot.title = element_text(face = "bold", hjust = 0.5))
}
```
```{r}
# Determine top 8 benefits based on overall selection %
top8 <- overall_ben %>%
  arrange(desc(pct)) %>%
  slice_head(n = 8) %>%
  pull(benefit)

# Apply readable labels (extend if needed)
benefit_labels <- c(
  "benb_mp01"  = "Improved Job Matching",
  "benb_mp02"  = "More Efficient Hiring",
  "benb_rvc03" = "Reduced Human Bias",
  "benb_can01" = "Career Advancement",
  "benb_ps01"  = "Personal Safety",
  "benb_vass02"= "Accessible Services",
  "benb_bc01"  = "Better Customer Service",
  "benb_rvc01" = "Reduced Cost"
)

plot_df <- gender_benefit %>%
  filter(benefit %in% top8) %>%
  mutate(
    benefit_label = recode(benefit, !!!benefit_labels, .default = benefit),
    benefit_label = factor(
      benefit_label,
      levels = recode(top8, !!!benefit_labels, .default = top8)
    )
  )

ggplot(plot_df, aes(x = benefit_label, y = pct, fill = sex)) +
  geom_col(position = "dodge") +
  coord_flip() +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_fill_manual(values = c("Female" = "pink", "Male" = "lightblue")) +
  labs(
    title = "Perceived Benefits of AI in Hiring by Gender",
    x = NULL,
    y = "Unweighted % selecting",
    fill = "Gender"
  ) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))



```



The chart compares how men and women perceive AI’s key benefits. Both
groups highlight Reduced Human Bias and Career Advancement as top
advantages, but women tend to rate Personal Safety and Accessible
Services slightly higher. These patterns suggest that while perceptions
of AI are largely similar across genders, women may place greater
emphasis on fairness and safety-related benefits.

Age vs Number of Benefits Seleced

```{r}
ai_clean <- ai_clean %>%
  mutate(benefit_count = rowSums(select(., all_of(benb_cols)) == 1, na.rm = TRUE))

ggplot(ai_clean, aes(x = age, y = benefit_count)) +
  geom_point(alpha = 0.3, color = "steelblue") +
  geom_smooth(method = "loess", color = "darkorange", se = TRUE) +
  labs(
    title = "Relationship Between Age and Number of AI Benefits Selected",
    x = "Age (Years)",
    y = "Number of Benefits Selected"
  ) +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(face = "bold", hjust = 0.5))
```

The scatter plot shows that younger respondents tend to identify more AI
benefits overall, while the number of perceived benefits declines
gradually with age. This suggests that comfort and optimism toward AI
may decrease with age.

## Conclusion

## Next Steps

