---
title: "Perceptions of Classical AI Across Demographics"
author: "Richard Liu, Eileena Mathews, Rachel Perry, Hayden Pozzi"
date: "`r Sys.Date()`"
output: html_document
---


```{r setup, include = FALSE}
# Load all required packages
library(ggplot2)
library(dplyr)
library(forcats)
library(tidyr)
library(stringr)
library(scales)
library(readr)
library(janitor)
library(corrplot)

# Global knit options
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
set.seed(2025)

```

```{r, include=FALSE}
library(scales)
# setup: load libraries and options
library(tidyverse)
library(readr)
library(haven)
library(janitor)
library(scales)
library(stringr)
library(dplyr)
library(tidyr)


knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
set.seed(2025)
```

# Introduction

In recent years, artificial intelligence has become increasingly embedded in everyday life, shaping how people interact with services, information, and technology. This project examines how the public perceived these systems using data collected in late 2022—immediately before generative AI tools like ChatGPT entered mainstream awareness. Because this survey predates the recent surge in AI hype, it provides a valuable baseline for attitudes toward classical, non-generative AI systems that people were actually encountering at the time (e.g., facial recognition, automated decision tools, and robotic care assistants).

The dataset we use is the **Public Attitudes to AI** survey from the Ada Lovelace Institute, which includes more than 4,000 UK respondents and 715 structured variables. It contains detailed measures of demographics, AI awareness, perceived benefits, and concerns across multiple domains. Since classical and generative AI are often conflated today, the timing of this dataset allows us to isolate perceptions of classical AI before generative models began reshaping public understanding.

Early exploratory plots in our analysis suggested meaningful demographic patterns—particularly by age, gender, and education—and these differences appeared especially pronounced in accessibility-related technologies such as Robotic Care Assistants (RCAs). These initial observations motivated the central question guiding our work:

**How do demographic characteristics shape public attitudes toward classical AI systems?**

To answer this, we combine descriptive summaries with regression models to test which demographic factors meaningfully predict perceived benefits or concerns. This structure allows us to move from broad patterns to statistical evidence, highlighting how different groups evaluate emerging AI technologies in a pre-LLM context.


## The Dataset:Public Attitudes to AI

```{r include=FALSE}
## Loading the raw data set
ai_path <- "how-do-people-feel-about-ai-main/262400371_Public Attitudes to AI Raw Data NA (1).csv"
if (!file.exists(ai_path)) stop("Data file not found: ", ai_path)
ai_survey1 <- read_csv(ai_path, show_col_types = FALSE)
head(ai_survey1, 10)
```

**Understanding the variables**
The dataset includes demographic variables such as age, sex, and region, along with a large set of binary columns prefixed with “benb_” that represent whether respondents selected specific perceived benefits of AI. Here we highlight the key variables used throughout our analysis. This table summarizes the most important features in our preprocessed dataset and provides the foundation for the models and visualizations presented below.

```{r echo=FALSE}
tibble::tribble(
~variable,       ~description,                                    ~type,
"age",           "Respondent age in years",                       "Numeric",
"sex",           "Self-identified gender (Male/Female/Other)",    "Categorical",
"benb_*",        "Binary indicators for perceived AI benefits",   "Integer (0/1)",
"weight",        "Survey weight (if applicable)",                 "Numeric"
) %>%
knitr::kable(caption = "Key variables in the preprocessed dataset")
```

## Tidying up ##
We preform additional tidyding such as recoding technology labels to match natural language,
listing demographic variables of interest, identifying and preparing key column key and creating "ai_clean" data set, standardizing variable names, detecting age and gender columns,
identifying all AI benefit-related columns (benb\_), and inspecting variables and "N/A" values.

However, for the sake of keeping our document conscise and informative, we have decided to omit
these steps from our report. To view our further tidying, visit the associated **.Rmd** file.

Likewise, we will artfully omit code from our analysis. To use an engineering analogy, this is to concentrate the reader on the "signal" rather than the "noise".

<!-- [*Recoding technology labels*]{.underline} -->

```{r echo=FALSE}
recode_technologies <- function(Technology_specific) {
  recode(Technology_specific,
                  "WEL" = "Welfare eligibility",
                "JOB" =  "Job eligibility",
                "CAN" =  "Cancer risk",
                "LOAN" = "Loan repayment risk", 
                "MP" =  "Mobile phone unlock",
                 "BC" = "Border control",
                 "PS" = "Police surveillance",

                "AW"  = "Autonomous weapons", 
                "DC" =  "Driverless cars",
                "RCA" = "Robotic care assistant",
               "RVC" ="Robotic vacuum cleaner",
                                                                         
               "SMC"=  "Targeted consumer ads",
                "SMP" = "Targeted political ads",
                                                                         
                "VASS"=  "Smart speaker",
                "VAH"=  "Virtual healthcare assistant",
                                 
              "CCR"=   "Climate change simulation",
                "VR"=  "Education simulations")
  
}
```

<!-- [*List of demographic variables of interest*]{.underline} -->

```{r echo=FALSE}
demog <- ai_survey1 %>%
  haven::zap_labels() %>%
  dplyr::select(
  Respondent_ID, PV25_Weight,
  RS_Ethnicity_Reduced,
  CS_Age_Band,
  RS_Sex,
  RS_EducationLevel,
  DR_Socio_Eco
  ) %>%
  mutate(RS_Ethnicity_Combined = 
           dplyr::recode(RS_Ethnicity_Reduced,
             "1" = "White",
             "2" = "Asian or Asian British",
             "3" = "Black British, Caribbean or African",
             "4" = "Any other ethnic group"
           )) %>%
  mutate(CS_Age_Band = 
           dplyr::recode(CS_Age_Band,
             "1" = "18-24 yrs",
             "2" = "25-34 yrs",
             "3" = "35-44 yrs",
             "4" = "45-54 yrs",
             "5" = "55-64 yrs",
             "6" = "65-74 yrs",
             "7" = "75+ yrs"
           )) %>%
  mutate(RS_Sex = 
           dplyr::recode(RS_Sex,
             "1" = "Male",
             "2" = "Female")) %>%
  mutate(RS_EducationLevel = 
           dplyr::recode(RS_EducationLevel,
             "1" = "Degree level qualification(s)",
             "2" = "Non-degree level qualifications",
             "3" = "No academic or vocational qualifications")) %>%
  mutate(DR_Socio_Eco = 
           dplyr::recode(DR_Socio_Eco,
             "1" = "SEC1, 2",
             "2" = "SEC3",
             "3" = "SEC4",
             "4" = "SEC5",
             "5" = "SEC6, 7",
             "6" = "SEC8",
             "7" = "Students")) %>%
  pivot_longer(names_to = "Q_Demo",
               values_to = "A_Demo",
               RS_Ethnicity_Combined:DR_Socio_Eco)

# list of levels 
demo_list <- c("White",
  "Asian or Asian British",
  "Black British, Caribbean or African", 
  "Any other ethnic group", 
  
  "18-24 yrs",
  "25-34 yrs",
  "35-44 yrs",
  "45-54 yrs",
  "55-64 yrs",
  "65-74 yrs",
  "75+ yrs", 
  
  "Male",
  "Female",
  "Identify in another way",
  
  "Degree level qualification(s)",
  "Non-degree level qualifications",
  "No academic or vocational qualifications",
  
  "SEC1, 2",
  "SEC3",
  "SEC4",
  "SEC5",
  "SEC6, 7",
  "SEC8",
  "Students", 
  
  "Total" )

if ("age" %in% names(ai_survey1)) {
ggplot(filter(ai, !is.na(age)), aes(age)) +
geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
labs(title = "Age Distribution of Respondents", x = "Age (years)", y = "Count")
}

```

<!-- [*Identify and prepare key column key and creating "ai_clean" dataset*]{.underline} -->

<!-- We standardize variable names, detect age and gender columns and
identify all AI benefit-related columns (benb\_). -->

```{r echo=FALSE}
## Basic cleaning and derived columns
ai <- read_csv(ai_path, show_col_types = FALSE)
# Standardise names and add row id
ai_clean <- ai_survey1 %>%
  clean_names() %>%
  mutate(row_id = row_number())

ai_clean <- ai %>%
clean_names() %>%
mutate(row_id = row_number())

# Detect age column

age_col <- names(ai_clean)[str_detect(names(ai_clean), regex("age", ignore_case = TRUE))][1]
# Detect age column (common names cs_age, csage, cs_age_yrs, age)
age_col <- names(ai_clean)[str_detect(names(ai_clean), regex("\\b(cs_age|csage|age|age_years|cs_age_yrs)\\b", ignore_case = TRUE))] %>% .[1]
if (!is.na(age_col)) {
ai_clean <- ai_clean %>%
mutate(age = as.integer(.data[[age_col]])) %>%
filter(is.na(age) | age >= 17)
  ai_clean <- ai_clean %>%
    mutate(age = as.integer(.data[[age_col]])) %>%
    filter(is.na(age) | age >= 17)  # remove implausible 16 if present
}

# Detect sex/gender column

sex_col <- names(ai_clean)[str_detect(names(ai_clean), regex("sex|gender", ignore_case = TRUE))][1]
sex_col <- names(ai_clean)[str_detect(names(ai_clean), regex("^rs_sex$|sex|gender", ignore_case = TRUE))] %>% .[1]
if (!is.na(sex_col)) {
ai_clean <- ai_clean %>%
mutate(
sex_raw = as.character(.data[[sex_col]]),
sex = case_when(
sex_raw %in% c("1", "Male", "male", "M") ~ "Male",
sex_raw %in% c("2", "Female", "female", "F") ~ "Female",
TRUE ~ NA_character_
)
)
  ai_clean <- ai_clean %>%
    mutate(
      sex_raw = as.character(.data[[sex_col]]),
      sex = case_when(
        sex_raw %in% c("1", 1, "Male", "male", "M") ~ "Male",
        sex_raw %in% c("2", 2, "Female", "female", "F") ~ "Female",
        TRUE ~ NA_character_
      )
    )
}

# Detect benefit-related columns

benb_cols <- names(ai_clean)[str_detect(names(ai_clean), regex("^benb|benefit", ignore_case = TRUE))]
# glimpse(select(ai_clean, any_of(c("age", "sex", benb_cols))))

```

<!-- [*Inspect variables and "missingness"*]{.underline} -->

## Exploring Missingness ##
Before getting deeper into the data, we wanted to check for missing
variable types to understand parts of the questionnaire that were left
unanswered by parts of the survey population. This would give insight
into variables that are interesting, but that might have missing context
because of their lack of completion by survey respondents. We will not be using these in our analysis.

```{r echo=FALSE}
name_map <- c(
  "governance_9" = "AI Governance – Who should ensure AI is used safely?",
  "benb_loan_7"  = "AI & Loans (Benefits) – How can AI loan risk prediction be beneficial?",
  "benb_rvc_4"   = "Robotic Vacuums (Benefits) – How are robotic vacuums beneficial?",
  "conb_loan_12" = "AI & Loans (Concerns) – What are your concerns about AI loan risk prediction?",
  "benb_aw_7"    = "Autonomous Weapons (Benefits) – How could autonomous weapons be beneficial?",
  "benb_ps_6"    = "Face Recognition in Policing (Benefits) – How could it benefit policing/surveillance?",
  "benb_bc_7"    = "Face Recognition at Borders (Benefits) – How could it benefit border control?",
  "benb_ccr_5"   = "AI & Climate Research (Benefits) – How could AI simulations help climate research?",
  "benb_can_6"   = "AI & Cancer Prediction (Benefits) – How could AI cancer prediction be beneficial?",
  "benb_smc_4"   = "Targeted Ads (Benefits) – How could targeted social media ads be beneficial?"
)
ai_missing <- tibble( variable = names(ai_clean), missing = map_int(ai_clean, ~ sum(is.na(.))), missing_pct = round(map_dbl(ai_clean, ~ mean(is.na(.))) * 100, 2) ) %>% arrange(desc(missing_pct)) %>% slice_head(n = 10)

ai_missing_disp <- ai_missing %>%
  mutate(
    variable_display = dplyr::recode(variable, !!!name_map, .default = variable)
    # If you want to keep the code too, use:
    # variable_display = dplyr::recode(variable, !!!name_map, .default = variable) |>
    #                    paste0(" (", variable, ")")
  ) %>%
  select(variable = variable_display, missing, missing_pct)

knitr::kable(ai_missing_disp, caption = "Top 10 Variables with the Highest Missingness")

# Detect BENB-like columns and coerce to 0/1 integers where possible
benb_cols <- names(ai_clean)[str_detect(names(ai_clean), regex("^benb|benb_job|benefit|benefits|benb_", ignore_case = TRUE))]
if (length(benb_cols) == 0) {
  benb_cols <- names(ai_clean)[str_detect(names(ai_clean), regex("benefit|benb|job", ignore_case = TRUE))]
}

if (length(benb_cols) > 0) {
  ai_clean <- ai_clean %>%
    mutate(across(all_of(benb_cols), ~ {
      x <- .
      # character encodings -> 1/0
      if (is.character(x)) {
        xl <- str_to_lower(str_trim(x))
        as.integer(if_else(xl %in% c("yes","y","1","selected","true","t"), 1L,
                   if_else(xl %in% c("no","n","0","false","f"), 0L, NA_integer_)))
      } else if (is.logical(x)) {
        as.integer(x)
      } else {
        # numeric-like: keep 1/0, else NA
        xi <- suppressWarnings(as.integer(x))
        if_else(is.na(xi), NA_integer_, xi)
      }
    }))
}


```

<!-- Summary Statistics -->
```{r summary-stats}
# Age summaries
if ("age" %in% names(ai_clean)) {
ai_clean %>%
summarise(
mean_age = mean(age, na.rm = TRUE),
median_age = median(age, na.rm = TRUE),
sd_age = sd(age, na.rm = TRUE),
min_age = min(age, na.rm = TRUE),
max_age = max(age, na.rm = TRUE)
)
  age_stats <- ai_clean %>%
    summarise(
      n = sum(!is.na(age)),
      mean_age = mean(age, na.rm = TRUE),
      median_age = median(age, na.rm = TRUE),
      sd_age = sd(age, na.rm = TRUE),
      min_age = min(age, na.rm = TRUE),
      max_age = max(age, na.rm = TRUE)
    )
  print(age_stats)
}

# Sex / gender distribution
if ("sex" %in% names(ai_clean)) {
ai_clean %>%
count(sex, sort = TRUE) %>%
mutate(percent = scales::percent(n / sum(n)))
  print(ai_clean %>% count(sex, sort = TRUE) %>% mutate(p = n / sum(n)))
}
```

## Visualizations and findings

[**Age Histogram**]{.underline}

```{r plots, echo=FALSE, fig.height=4, fig.width=7}
# Age histogram
if ("age" %in% names(ai_clean)) {
ggplot(filter(ai_clean, !is.na(age)), aes(age)) +
geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
labs(title = "Age Distribution", x = "Age", y = "Count")
  ggplot(filter(ai_clean, !is.na(age)), aes(age)) +
    geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
    labs(title = "Age distribution", x = "Age", y = "Count")
}

```

Because one of our core research questions examines how age shapes attitudes toward classical AI systems, confirming that we have a broad age range was essential. This distribution reassures us that the data include enough representation across younger, middle-aged, and older adults to meaningfully explore how attitudes may differ by age. The relatively smaller number of respondents over 70 is important to keep in mind, but the overall spread still provides a strong foundation for testing age-related patterns that appear in our later descriptive plots and regression models.

[**Gender Distribution**]{.underline}

```{r echo=FALSE}
# Gender bar chart

# Sex / gender bar
if ("sex" %in% names(ai_clean)) {
ggplot(filter(ai_clean, !is.na(sex)), aes(x = fct_infreq(sex))) +
geom_bar(fill = "darkorange") +
coord_flip() +
labs(title = "Gender Distribution", x = NULL, y = "Count")
  ggplot(filter(ai_clean, !is.na(sex)), aes(x = fct_infreq(sex))) +
    geom_bar(fill = "darkorange") +
    coord_flip() +
    labs(title = "Sex / Gender distribution", x = NULL, y = "Count")
}
```

The chart shows that survey responses were evenly split between male and
female participants, giving a balanced view of public attitudes toward
AI. This balance helps ensure that later comparisons by gender aren’t
biased by unequal sample sizes.

[*Top 8 Perceived Benefits of AI*]{.underline}

**Derive Overall Benefit Tallies (Unweighted)**

```{r derive-overall-ben, echo=FALSE}
overall_ben <- ai_clean %>%
  tidyr::pivot_longer(all_of(benb_cols), names_to = "benefit", values_to = "selected",
                      values_transform = list(selected = as.integer),
                      values_drop_na = TRUE) %>%
  dplyr::filter(selected == 1L) %>%
  dplyr::count(benefit, sort = TRUE) %>%
  dplyr::mutate(pct = n / sum(n))
```

```{r echo=FALSE}
overall_ben <- overall_ben %>%
  mutate(
    benefit = str_to_lower(str_trim(benefit)),  # clean names first
    benefit_label = recode(benefit,
      "benb_mp01" = "Faster to unlock devices",
      "benb_mp02" = "Personal information more safe and secure",
      "benb_rvc03" = "Vacuuming possible for people who have trouble with manual tasks",
      "benb_can01" = "Earlier detection of cancer",
      "benb_ps01"  = "Identify wanted criminals and missing persons",
      "benb_vass02" = "People with difficulty using devices have better access to features",
      "benb_bc01"  = "Processing people at border control will be faster",
      "benb_rvc01" = "Tech can do vacuuming and save people time",
      .default = benefit  # keeps unmapped names so nothing breaks
    )
  )


# Select Top 8 by count
top8 <- overall_ben %>%
  arrange(desc(n)) %>%
  slice_head(n = 8) %>%
  mutate(
    benefit_label = factor(
      benefit_label,
      levels = benefit_label[order(n, decreasing = TRUE)]
    )
  )
```

```{r echo=FALSE}
# Plot Top 8 with recoded labels
ggplot(top8, aes(x = n, y = benefit_label)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = n), hjust = -0.1, size = 3.7) +
  coord_cartesian(xlim = c(0, max(top8$n) * 1.15)) +
  labs(
    title = "Top 8 Benefits Selected",
    subtitle = "Raw number of respondents selecting each benefit",
    x = "Number of Respondents",
    y = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    axis.text.y = element_text(size = 10)
  )

```

The raw response data suggests that there is no single dominant theme in how the public perceives the benefits of AI. Instead, respondents point to a wide range of everyday advantages—from faster device access to improved personal information security, enhanced accessibility for people with difficulty using devices, and support for public safety and border processes. These diverse selections show that people primarily associate AI with practical, non-generative tools that are already embedded in daily routines.

This broad spread of uses reinforced our interest in studying classical (non-generative) AI, since its applications appear more widespread and familiar to the public than any single high-profile technology. Because perceived benefits vary so widely, this figure also motivated us to examine whether demographic characteristics—such as age, gender, or education—shape which advantages individuals recognize.

## Looking deeper - AI Usage for Accessibility Purposes
Because the perceived benefits of AI span a variety of domains, we next focused our analysis on a specific application to better understand how attitudes differ across demographic groups. We selected Robo-vacuums and robotic care assistants (RCAs) as a case example: everyday forms of classical AI that can meaningfully improve accessibility, particularly for individuals with mobility limitations or other disabilities. Prior research has noted that assistive cleaning technologies can reduce physical strain and support independent living among vulnerable populations*([Cleaning Is
Caring,
n.d.)](https://www.cleaningiscaring.org/wpd/disability-friendly-cleaning/)*.

RCAs are especially interesting in the context of our project because, as highlighted in our presentation, awareness of these technologies is relatively similar across age groups, even though RCAs are often designed to support older adults or people who benefit from additional assistance in daily tasks. This combination—broad, cross-age familiarity paired with potentially age-specific usefulness—makes RCAs a compelling domain for testing whether demographic characteristics such as age, gender, and education shape how people perceive accessibility-related benefits of AI.

By examining responses related to Robo-vacuum and RCA technologies, we investigate whether different demographic groups recognize these benefits to the same extent, and whether age plays a role in shaping attitudes toward assistive AI systems.
)*.
By examining responses related to Robo-vacuum and robo-assistants , we
explore whether age, gender, and other demographic characteristics
influence who recognizes these accessibility-related benefits of AI.

```{r echo=FALSE}
# Identify column name automatically if needed
aware_var <- names(ai_clean)[str_detect(names(ai_clean), regex("aware_rca", ignore_case = TRUE))]
stopifnot(length(aware_var) == 1)

# Create age group bands if numeric age exists
age_group_df <- ai_clean %>%
  filter(!is.na(.data[[aware_var]])) %>%
  mutate(
    awareness = case_when(
      str_to_lower(str_trim(.data[[aware_var]])) %in% c("1","yes","y","true") ~ "Yes",
      str_to_lower(str_trim(.data[[aware_var]])) %in% c("0","no","n","false") ~ "No",
      TRUE ~ "Not sure"
    ),
    age_group = case_when(
      age < 25 ~ "18–24",
      age >= 25 & age <= 34 ~ "25–34",
      age >= 35 & age <= 44 ~ "35–44",
      age >= 45 & age <= 54 ~ "45–54",
      age >= 55 & age <= 64 ~ "55–64",
      age >= 65 ~ "65+",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(age_group))

aware_age <- ai_clean %>%
  select(age_group = age, aware_rca = aware_rca) %>%
  filter(!is.na(age_group), aware_rca %in% c(1,2,3)) %>%
  mutate(
    aware = case_when(
      aware_rca == 1 ~ 1L,        # Yes
      aware_rca %in% c(2,3) ~ 0L  # No + Not sure
    )
  ) %>%
  group_by(age_group) %>%
  summarise(pct_yes = mean(aware), .groups = "drop")

aware_age <- aware_age %>%
  mutate(age_num = as.numeric(age_group))  # Convert ordered factor to numeric for regression

ggplot(aware_age, aes(x = age_group, y = pct_yes)) +
  geom_col(fill = "#4B8EA8") +
   # Regression line
  geom_smooth(aes(x = age_num, y = pct_yes), 
              method = "lm", 
              se = FALSE, 
              color = "red", 
              size = 1,
              linetype = "dashed",
              inherit.aes = FALSE) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "Awareness of Robotic Care Assistants by Age",
    subtitle = "Percent who have heard of robotic care assistants",
    x = "Age Group", y = "% Aware"
  ) +
  theme_minimal(base_size = 14)
```

Although the bar heights suggest some fluctuation in awareness across ages, the overall pattern reveals that awareness of robotic care assistants is relatively low and does not meaningfully change with age. This means that despite RCAs being designed to support older adults and individuals who may benefit from additional assistance, familiarity with these technologies is not concentrated in any particular age group. Instead, most respondents—regardless of age—report limited awareness.

The subtle downward slope in the regression line likely reflects sampling variation, particularly because respondents over 75 make up only a small portion of the dataset, resulting in greater noise in the oldest age groups. Still, the broader takeaway remains clear: people’s attitudes toward RCAs are shaped in a context where many have heard little about them, making this domain especially useful for examining how demographic factors, rather than direct exposure, influence perceptions of AI designed for care and accessibility.

This pattern reinforced our choice to focus on RCAs as a case study. Their cross-age relevance, paired with low and relatively uniform awareness, makes them a compelling technology for investigating how age, gender, education, and other demographic predictors shape perceptions of assistive AI.

```{r include=FALSE}
ai_clean %>%
  summarise(
    total = n(),
    n_75plus = sum(age >= 75, na.rm = TRUE),
    pct_75plus = round(n_75plus / total * 100, 1)
  )

```

```{r echo=FALSE}
# Recode response values to labels
ai_plot <- ai_clean %>%
  select(age, bena_rca) %>%
  mutate(
    bena_rca = recode(as.character(bena_rca),
      "1" = "Very beneficial",
      "2" = "Fairly beneficial",
      "3" = "Not very beneficial",
      "4" = "Not at all beneficial",
      "5" = "Don't know",
      .default = NA_character_
    ),
    age_band = cut(
      age,
      breaks = c(17, 24, 34, 44, 54, 64, 74, Inf),
      labels = c("18–24", "25–34", "35–44", "45–54", "55–64", "65–74", "75+"),
      right = TRUE
    )
  ) %>%
  filter(!is.na(age_band), !is.na(bena_rca))

# Make simple bar chart
ggplot(ai_plot, aes(x = age_band, fill = bena_rca)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_brewer(palette = "Blues", direction = 1) +
  labs(
    title = "Perceived Benefit of Robotic Care Assistants by Age Group",
    subtitle = "Question: To what extent do you think the use of robotic care assistants will be beneficial?",
    x = "Age group",
    y = "Percent of respondents",
    fill = "Response"
  ) +
  theme_minimal(base_size = 12)
```
Across all age groups, most respondents believe that robotic care assistants would be at least “fairly beneficial,” indicating broad, baseline positivity toward this type of assistive technology. However, the distribution of responses also reveals subtle demographic patterns consistent with what we highlighted in our presentation: older adults appear somewhat less likely to select the highest benefit categories, while younger and middle-aged respondents show slightly stronger enthusiasm.

At the same time, “don’t know” responses remain relatively common across all age bands, reinforcing our earlier point that many people are forming opinions about RCAs without extensive familiarity. 

Taken together, this figure suggests that while attitudes toward RCAs are generally positive, age may still play a modest role in shaping perceived benefit. This descriptive pattern motivated us to test these differences more formally through regression modeling in the next section.

```{r echo=FALSE}
concern_age <- ai_clean %>%
  select(age_group = age, concern = cona_rca) %>% 
  filter(!is.na(age_group) & concern %in% c(1,2,3,4,5)) %>%
  mutate(
    concerned = case_when(
      concern %in% c(1,2) ~ 1L,   # Very + Somewhat concerned
      concern %in% c(3,4,5) ~ 0L  # Not concerned + Don't know
    )
  ) %>%
  group_by(age_group) %>%
  summarise(
    pct_concern = mean(concerned, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

ggplot(concern_age, aes(x = age_group, y = pct_concern, group = 1)) +
  
  geom_col(fill = "#D95F02", alpha = 0.7) +

  # Regression line
  geom_smooth(aes(x = as.numeric(age_group), y = pct_concern),
              method = "lm", se = FALSE,
              color = "black", size = 1,
              linetype = "dashed",
              inherit.aes = FALSE) +
  
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0,1)) +
  
  labs(
    title = "Concern About Robotic Care Assistants by Age",
    subtitle = "Share of respondents who are very or somewhat concerned",
    x = "Age Group",
    y = "% Concerned"
  ) +

  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    panel.grid.minor = element_blank()
  )

```

Concern about robotic care assistants shows a clear upward trend with age, suggesting that older adults are somewhat more wary of this technology. This aligns with patterns highlighted in our presentation, where age emerged as a meaningful predictor of caution toward assistive AI. However, the bar heights also reveal that elevated concern is present across all age groups, not just older respondents. Even younger adults show substantial levels of concern, indicating that unease around RCAs is a broadly shared sentiment rather than an older-adult phenomenon.

```{r echo=FALSE}
# Select all BENB_DC columns
benb_dc_data <- ai_clean %>%
  select(starts_with("benb_dc"))

# Convert to long format
benb_dc_long <- benb_dc_data %>%
  mutate(id = row_number()) %>%  # respondent ID
  pivot_longer(
    cols = starts_with("benb_dc"),
    names_to = "option",
    values_to = "selected"
  )

# Summarise counts and percentages
benb_dc_summary <- benb_dc_long %>%
  filter(!is.na(selected)) %>%
  group_by(option) %>%
  summarise(
    count = sum(selected == 1, na.rm = TRUE),
    total_respondents = n_distinct(id),
    percent = (count / total_respondents) * 100
  ) %>%
  ungroup() %>%
  # Clean up inconsistent naming
  mutate(
    option = tolower(option),
    option = str_replace_all(option, "benb[_]?dc[_]?", "benb_dc"),
    option = str_replace_all(option, "0+", "")
  ) %>%
  # Remove unwanted items
  filter(!option %in% c("benb_dc10", "benb_dc11")) %>%
  arrange(desc(percent))

# ---- PLOT ----
ggplot(benb_dc_summary, aes(x = reorder(option, percent), y = percent)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  scale_x_discrete(
    labels = c(
      "benb_dc1" = "Easier car travel",
      "benb_dc2" = "More time for other activities",
      "benb_dc3" = "More accurate driving",
      "benb_dc4" = "Fewer accidents",
      "benb_dc5" = "Easier for disabled or older drivers",
      "benb_dc6" = "Saves money on drivers",
      "benb_dc7" = "Something else",
      "benb_dc8" = "None of these",
      "benb_dc9" = "Don't know"
    )
  ) +
  labs(
    title = "Perceived Benefits of Driverless Cars",
    x = NULL,
    y = "Percentage of Respondents (%)"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10),
    plot.title = element_text(face = "bold", size = 14)
  )
```

Respondents highlight a range of potential advantages of driverless cars, but the most commonly selected benefit is their ability to make driving easier for disabled or older adults. This aligns with the broader pattern we observed in our analysis: many perceived benefits of classical AI technologies relate to accessibility, reduced physical strain, and increased independence for vulnerable populations. Safety-related benefits, such as fewer accidents and more accurate driving, also rank highly, suggesting that respondents associate driverless cars with both personal convenience and public safety.

Notably, benefits like saving money or having more time for other activities are selected less frequently, indicating that respondents tend to focus more on practical, everyday improvements rather than abstract or economic advantages. The relatively small “don’t know” proportion further suggests that people feel comfortable expressing an opinion about this technology, even if they do not have direct experience with driverless cars, similar to what we saw with other RCAs.

## Exploring Benefit Score vs. Age: Bridge to Modeling
```{r echo=FALSE}
if (length(benb_cols) > 0 && "age" %in% names(ai_clean)) {

  ai_model_df <- ai_clean %>%
    mutate(
      benefit_score = rowSums(across(all_of(benb_cols)), na.rm = TRUE)
    ) %>%
    select(age, benefit_score) %>%
    filter(!is.na(age), !is.na(benefit_score))

  ggplot(ai_model_df, aes(x = age, y = benefit_score)) +
    geom_point(alpha = 0.2) +
    geom_smooth(method = "lm", se = FALSE, color = "red") +
    labs(
      title = "Total Number of Benefits Selected vs Age",
      subtitle = "A simple illustration of how a regression model may treat benefit score as an outcome",
      x = "Age",
      y = "Benefit score (count of benefits selected)"
    ) +
    theme_minimal()
}
```
The scatterplot shows that although people across all ages select a wide range of AI benefits, there is no strong linear relationship between age and the total number of benefits selected. The regression line is almost flat, suggesting that older adults are not necessarily more or less likely than younger adults to see AI as beneficial overall.

However, the large spread in benefit scores at every age indicates that age alone cannot explain variation in attitudes. This reinforces why a regression model is needed: to test how age interacts with other factors such as gender, education, socio-economic status, and awareness—to more accurately predict who perceives AI as beneficial.

```{r include=FALSE}
## Prepare data for regression

# 1. Overall benefit score: sum across all benb_ columns
ai_clean$model_benefit_score <- rowSums(ai_clean[, benb_cols], na.rm = TRUE)

# 2. Recode outcomes for Robotic Care Assistants (RCA)
ai_clean <- ai_clean %>%
  mutate(
    rca_beneficial = case_when(
      bena_rca %in% c(1, 2) ~ 1,   # Very or fairly beneficial
      bena_rca %in% c(3, 4) ~ 0,   # Not very / not at all beneficial
      TRUE ~ NA_real_
    ),
    rca_concerned = case_when(
      cona_rca %in% c(1, 2) ~ 1,   # Very or somewhat concerned
      cona_rca %in% c(3, 4, 5) ~ 0, # Not very / not at all / don't know
      TRUE ~ NA_real_
    )
  )

# 3. Build modeling dataset with the correct column names
model_df <- ai_clean %>%
  select(
    age,
    sex,
    rs_education_level,  # <-- this is the real name in your data
    dr_socio_eco,        # <-- socio-economic group
    model_benefit_score,
    rca_beneficial,
    rca_concerned
  ) %>%
  filter(!is.na(age), !is.na(sex))

head(model_df)


```
```{r include=FALSE}
# Convert categorical predictors to factors
model_df <- model_df %>%
  mutate(
    sex = factor(sex),
    rs_education_level = factor(rs_education_level,
                                labels = c("Degree", "Non-degree", "No qualifications")),
    dr_socio_eco = factor(dr_socio_eco)
  )
```

## Model 1: Benefit Score ~ Age

```{r echo=FALSE}
model1 <- lm(model_benefit_score ~ age, data = model_df)
summary(model1)
```

Age is not a significant predictor of the total number of AI benefits a respondent selects. The estimated slope is very close to zero, and the associated p-value is far too large to suggest any meaningful relationship between age and benefit scores. In fact, the model explains almost none of the variation in perceived benefits. This aligns with our earlier scatterplot: people across all age groups report similar levels of perceived benefit, indicating that age alone does not drive how many advantages individuals attribute to AI.

## Model 2: Logistic Regression (Who thinks RCAs are beneficial?)

```{r echo=FALSE}
model2 <- glm(
  rca_beneficial ~ age + sex + rs_education_level + dr_socio_eco,
  data = model_df,
  family = binomial()
)
summary(model2)
```
In this model, several demographic variables significantly predict whether respondents view robotic care assistants as beneficial.

First, **age has a small but statistically significant negative association** with perceived benefit (p < 0.001). Although the effect size is modest, the negative coefficient indicates that **older adults are slightly less likely** to view RCAs positively—an insight consistent with our earlier descriptive findings.

Gender also plays a meaningful role: **men are significantly more likely than women** to view RCAs as beneficial (p < 0.001), echoing the stronger enthusiasm for assistive technologies observed in our exploratory plots.

Education shows a clear gradient. Compared to respondents with degree-level qualifications, those with **non-degree qualifications or no formal qualifications** are significantly less likely to view RCAs as beneficial (both p < 0.01). This suggests that familiarity with technology or higher educational attainment may shape comfort with or perceived usefulness of AI in care settings.

Socioeconomic classification exhibits smaller and more mixed effects—some SEC groups show reduced odds of perceiving RCAs as beneficial, though these effects are less consistent across categories.

Overall, this model confirms the descriptive patterns we highlighted earlier: **men and more highly educated respondents are more likely to see RCAs as beneficial, while older adults tend to be slightly more cautious**. These results reinforce the idea that demographic factors shape perceptions of classical, non-generative AI technologies in meaningful ways.
```{r coef_plot_model2, echo=FALSE, fig.height=5, fig.width=7}
# Model 2: graph
m2_tidy <- broom::tidy(model2, conf.int = TRUE, exponentiate = TRUE) %>%
  dplyr::filter(term != "(Intercept)") %>%
  dplyr::mutate(
    term = dplyr::recode(term,
                         "sexMale" = "Male (vs Female)",
                         "age" = "Age (per year)",
                         "rs_education_levelNon-degree" = "Non-degree (vs Degree)",
                         "rs_education_levelNo qualifications" = "No qualifications (vs Degree)",
                         "dr_socio_eco2" = "SEC 2",
                         "dr_socio_eco3" = "SEC 3",
                         "dr_socio_eco4" = "SEC 4",
                         "dr_socio_eco5" = "SEC 5",
                         "dr_socio_eco6" = "SEC 6",
                         "dr_socio_eco7" = "SEC 7"
                        ),
    term = forcats::fct_reorder(term, estimate)
  )

ggplot(m2_tidy, aes(x = term, y = estimate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.1) +
  coord_flip() +
  labs(
    title = "Odds Ratios for Viewing RCAs as Beneficial",
    x = NULL,
    y = "Odds Ratio (95% CI)"
  ) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  theme_minimal()
```

<!-- ### Model 2: RCA Beneficial ~ Age + Sex + Education + SEC -->
The odds-ratio plot reinforces the regression results: men have substantially higher odds of viewing RCAs as beneficial, while respondents with lower educational attainment have meaningfully lower odds compared to those with degrees. Age shows a small but significant negative association, reflected by an odds ratio slightly below 1. Socioeconomic categories vary, but most hover near 1 with wide confidence intervals, indicating weaker or inconsistent effects. Overall, these patterns mirror the conclusions from Model 2 and visually highlight which demographic characteristics have the strongest influence on perceived benefit.

## Model 3: Logistic Regression (Who is concerned about RCAs?)

```{r include=FALSE}
model3 <- glm(
  rca_concerned ~ age + sex + rs_education_level + dr_socio_eco,
  data = model_df,
  family = binomial()
)
summary(model3)
```

```{r coef_plot_model3, echo=FALSE, fig.height=5, fig.width=7}
# Model 3: graph
m3_tidy <- broom::tidy(model3, conf.int = TRUE, exponentiate = TRUE) %>%
  dplyr::filter(term != "(Intercept)") %>%
  dplyr::mutate(
    term = dplyr::recode(
      term,
      "sexMale" = "Male (vs Female)",
      "age" = "Age (per year)",
      "rs_education_levelNon-degree" = "Non-degree (vs Degree)",
      "rs_education_levelNo qualifications" = "No qualifications (vs Degree)",
      "dr_socio_eco2" = "SEC 2",
      "dr_socio_eco3" = "SEC 3",
      "dr_socio_eco4" = "SEC 4",
      "dr_socio_eco5" = "SEC 5",
      "dr_socio_eco6" = "SEC 6",
      "dr_socio_eco7" = "SEC 7"
    ),
    term = forcats::fct_reorder(term, estimate)
  )

ggplot(m3_tidy, aes(x = term, y = estimate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.1) +
  coord_flip() +
  labs(
    title = "Odds Ratios for Concern About RCAs",
    x = NULL,
    y = "Odds Ratio (95% CI)"
  ) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  theme_minimal()
```


<!-- ### Model 3: RCA Concerned ~ Age + Sex + Education + SEC -->
Age is a significant predictor of concern about robotic care assistants. The positive coefficient for age (p < 0.001) indicates that **older respondents are more likely to report being somewhat or very concerned about RCAs**, which is consistent with the upward trend we saw in the concern-by-age figure. 

Gender shows the opposite pattern of the benefit model: the large negative coefficient for sexMale (p < 0.001) means that men are substantially less likely than women to be concerned about RCAs. In other words, **women are more likely to express worry about this technology**, even after controlling for age, education, and socioeconomic status.

By contrast to the benefist model, education level is not a strong predictor of concern once other variables are included. Neither non-degree qualifications nor having no formal qualifications differ significantly from degree-holders in their odds of being concerned. The socioeconomic status indicators (SEC categories) also hover near zero and are not statistically significant, suggesting that class-based differences in concern are relatively weak in this model.

Taken together, Model 3 shows that age and gender, rather than education or socioeconomic status, are the main demographic drivers of concern about RCAs: **older adults and women are more likely to be uneasy about this form of assistive AI**, even though awareness is relatively similar across age groups.

## Conclusion 

Our analysis of the Public Attitudes to AI dataset shows that perceptions of classical AI technologies—including robotic care assistants—vary across demographic groups in meaningful but nuanced ways. While descriptive patterns suggested broad familiarity gaps and mixed reactions across age groups, the regression models help clarify which characteristics actually shape attitudes once multiple factors are considered simultaneously.

Across the models, **age and gender consistently emerge as the strongest predictors**. Older adults are modestly less likely to view RCAs as beneficial and more likely to express concern about their use. Gender differences also appear in opposite directions across the two outcomes: men are significantly more likely to view RCAs as beneficial, yet substantially less likely to be concerned about them. These patterns suggest that demographic differences in comfort, familiarity, or trust may influence how people evaluate assistive AI technologies.

**Education and socioeconomic status, by contrast, play a more limited role**. In the benefit model, lower educational attainment reduces the likelihood of perceiving RCAs as useful, but these effects do not translate into heightened concern. Socioeconomic indicators show weak or inconsistent associations across both models, indicating that class-based differences are not central drivers of RCA perceptions in this sample.

Taken together, these findings highlight that **attitudes toward classical AI are shaped primarily by age and gender** rather than broad structural factors such as education level or socioeconomic status. Public responses to assistive technologies like RCAs are therefore influenced less by familiarity or technical background and more by demographic patterns in perceived utility, safety, and trust. Understanding these differences is essential for anticipating how various groups may engage with emerging AI applications, particularly those designed to support accessibility and care.

## Future Direction 

The models in this notebook focus on broad demographic predictors of how people view robotic care assistants. A natural next step would be to incorporate more direct measures of accessibility and disability—such as disability status, mobility limitations, and caring responsibilities—to better understand whether those who stand to benefit most from assistive technologies are more likely to view them as useful or concerning. Including measures of technology experience, confidence, and familiarity would also help clarify whether attitudes reflect practical exposure or broader feelings of trust in AI.

Methodologically, future work could explore interaction terms (for example, age × disability or gender × technology comfort) to test whether demographic groups respond differently depending on their everyday relationship with technology. Applying survey weights and more flexible modeling strategies—such as logistic regression with interactions, regularized models, or multilevel approaches—would allow these findings to generalize more reliably to the UK population.

There is also an important practical dimension: **as generative AI systems dominate public conversations about artificial intelligence, companies developing classical, non-generative AI systems have strong incentives to understand how their technologies are perceived**. Many longstanding applications of AI in healthcare, transportation, and accessibility have been overshadowed by narratives about ChatGPT and other generative tools. Mapping public attitudes toward non-generative AI can help these companies communicate their value, address sources of concern, and ensure that the benefits of assistive technologies are visible to the groups they are designed to support.