---
editor_options: 
  markdown: 
    wrap: 72

---

```{r setup, include = FALSE}
# Load all required packages
library(ggplot2)
library(dplyr)
library(forcats)
library(tidyr)
library(stringr)
library(scales)
library(readr)
library(janitor)
library(corrplot)

# Global knit options
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
set.seed(2025)

```

```{r, include=FALSE}
library(scales)
# setup: load libraries and options
library(tidyverse)
library(readr)
library(haven)
library(janitor)
library(scales)
library(stringr)
library(dplyr)
library(tidyr)


knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
set.seed(2025)
```

## Introduction

**Background**

In recent years, artificial intelligence (AI) has shifted from a source
of uncertainty and risk to a common feature of everyday life. From
writing assistants and chat bots to algorithmic recommendations in
education, finance, and entertainment, AI has become deeply embedded in
social and professional contexts. Understanding how people perceive its
benefits and risks is crucial for policymakers and organizations
designing AI systems that are equitable and trustworthy. In recent
years, artificial intelligence (AI) has undergone a profound shift in
public perception. Once seen primarily as a source of uncertainty or
even risk, AI is now becoming something people feel increasingly
comfortable with—if not socially or professionally expected to use. From
writing assistants and chatbots to algorithmic recommendations in
education, finance, and entertainment, AI has become an embedded feature
of everyday life.

**Dataset and preprocessing**

To ground our analysis, we use the Public Attitudes to AI (PAI) data set
produced by the Ada Lovelace Institute (UK). This nationally
representative survey of approximately 4,000 adults captures attitudes
toward AI across multiple domains—such as hiring, healthcare, and
policing. To ground our analysis, we use the Public Attitudes to AI
dataset produced by the Ada Lovelace Institute (UK). This nationally
representative survey of \~4,000 adults provides perceptions across
multiple AI application domains. Our aim here is a reproducible initial
exploration: inspect structure and missingness, perform light cleaning,
compute simple summaries, and visualise perceived benefits of AI in
hiring.

This project uses the preprocessed version of the data set provided by
the Ada Lovelace Institute. Our group utilized similar cleaning
techniques based on the descriptive analysis code by the Ada Lovelace
Institute.

**Goal:** Our project aims to replicate the results of the Ada Lovelace
Institution, but to also consider gender, age, and other demographics
that impacted survey respondents' perceived benefits of AI.

## Explaining the Dataset

**Loading raw data set**

```{r}
ai_path <- "how-do-people-feel-about-ai-main/262400371_Public Attitudes to AI Raw Data NA (1).csv"
if (!file.exists(ai_path)) stop("Data file not found: ", ai_path)
ai_survey1 <- read_csv(ai_path, show_col_types = FALSE)
head(ai_survey1, 10)
```

**Understanding the variables and tidying up**

The dataset includes demographic variables such as age, sex, and region
along with multiple binary columns (beginning with 'benb\_')
representing whether respondents selected each perceived benefit. Below
we summarize the key variables used in our analysis

```{r}
tibble::tribble(
~variable,       ~description,                                    ~type,
"age",           "Respondent age in years",                       "Numeric",
"sex",           "Self-identified gender (Male/Female/Other)",    "Categorical",
"benb_*",        "Binary indicators for perceived AI benefits",   "Integer (0/1)",
"weight",        "Survey weight (if applicable)",                 "Numeric"
) %>%
knitr::kable(caption = "Key variables in the preprocessed dataset")
```

[*Recoding technology labels*]{.underline}

```{r}
recode_technologies <- function(Technology_specific) {
  recode(Technology_specific,
                  "WEL" = "Welfare eligibility",
                "JOB" =  "Job eligibility",
                "CAN" =  "Cancer risk",
                "LOAN" = "Loan repayment risk", 
                "MP" =  "Mobile phone unlock",
                 "BC" = "Border control",
                 "PS" = "Police surveillance",

                "AW"  = "Autonomous weapons", 
                "DC" =  "Driverless cars",
                "RCA" = "Robotic care assistant",
               "RVC" ="Robotic vacuum cleaner",
                                                                         
               "SMC"=  "Targeted consumer ads",
                "SMP" = "Targeted political ads",
                                                                         
                "VASS"=  "Smart speaker",
                "VAH"=  "Virtual healthcare assistant",
                                 
              "CCR"=   "Climate change simulation",
                "VR"=  "Education simulations")
  
}
```

[*List of demographic variables of interest*]{.underline}

```{r}
demog <- ai_survey1 %>%
  haven::zap_labels() %>%
  dplyr::select(
  Respondent_ID, PV25_Weight,
  RS_Ethnicity_Reduced,
  CS_Age_Band,
  RS_Sex,
  RS_EducationLevel,
  DR_Socio_Eco
  ) %>%
  mutate(RS_Ethnicity_Combined = 
           dplyr::recode(RS_Ethnicity_Reduced,
             "1" = "White",
             "2" = "Asian or Asian British",
             "3" = "Black British, Caribbean or African",
             "4" = "Any other ethnic group"
           )) %>%
  mutate(CS_Age_Band = 
           dplyr::recode(CS_Age_Band,
             "1" = "18-24 yrs",
             "2" = "25-34 yrs",
             "3" = "35-44 yrs",
             "4" = "45-54 yrs",
             "5" = "55-64 yrs",
             "6" = "65-74 yrs",
             "7" = "75+ yrs"
           )) %>%
  mutate(RS_Sex = 
           dplyr::recode(RS_Sex,
             "1" = "Male",
             "2" = "Female")) %>%
  mutate(RS_EducationLevel = 
           dplyr::recode(RS_EducationLevel,
             "1" = "Degree level qualification(s)",
             "2" = "Non-degree level qualifications",
             "3" = "No academic or vocational qualifications")) %>%
  mutate(DR_Socio_Eco = 
           dplyr::recode(DR_Socio_Eco,
             "1" = "SEC1, 2",
             "2" = "SEC3",
             "3" = "SEC4",
             "4" = "SEC5",
             "5" = "SEC6, 7",
             "6" = "SEC8",
             "7" = "Students")) %>%
  pivot_longer(names_to = "Q_Demo",
               values_to = "A_Demo",
               RS_Ethnicity_Combined:DR_Socio_Eco)

# list of levels 
demo_list <- c("White",
  "Asian or Asian British",
  "Black British, Caribbean or African", 
  "Any other ethnic group", 
  
  "18-24 yrs",
  "25-34 yrs",
  "35-44 yrs",
  "45-54 yrs",
  "55-64 yrs",
  "65-74 yrs",
  "75+ yrs", 
  
  "Male",
  "Female",
  "Identify in another way",
  
  "Degree level qualification(s)",
  "Non-degree level qualifications",
  "No academic or vocational qualifications",
  
  "SEC1, 2",
  "SEC3",
  "SEC4",
  "SEC5",
  "SEC6, 7",
  "SEC8",
  "Students", 
  
  "Total" )

if ("age" %in% names(ai_survey1)) {
ggplot(filter(ai, !is.na(age)), aes(age)) +
geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
labs(title = "Age Distribution of Respondents", x = "Age (years)", y = "Count")
}

```

[*Identify and prepare key column key and creating "ai_clean"
dataset*]{.underline}

We standardize variable names, detect age and gender columns and
identify all AI benefit-related columns (benb\_).

```{r}
## Basic cleaning and derived columns
ai <- read_csv(ai_path, show_col_types = FALSE)
# Standardise names and add row id
ai_clean <- ai_survey1 %>%
  clean_names() %>%
  mutate(row_id = row_number())

ai_clean <- ai %>%
clean_names() %>%
mutate(row_id = row_number())

# Detect age column

age_col <- names(ai_clean)[str_detect(names(ai_clean), regex("age", ignore_case = TRUE))][1]
# Detect age column (common names cs_age, csage, cs_age_yrs, age)
age_col <- names(ai_clean)[str_detect(names(ai_clean), regex("\\b(cs_age|csage|age|age_years|cs_age_yrs)\\b", ignore_case = TRUE))] %>% .[1]
if (!is.na(age_col)) {
ai_clean <- ai_clean %>%
mutate(age = as.integer(.data[[age_col]])) %>%
filter(is.na(age) | age >= 17)
  ai_clean <- ai_clean %>%
    mutate(age = as.integer(.data[[age_col]])) %>%
    filter(is.na(age) | age >= 17)  # remove implausible 16 if present
}

# Detect sex/gender column

sex_col <- names(ai_clean)[str_detect(names(ai_clean), regex("sex|gender", ignore_case = TRUE))][1]
sex_col <- names(ai_clean)[str_detect(names(ai_clean), regex("^rs_sex$|sex|gender", ignore_case = TRUE))] %>% .[1]
if (!is.na(sex_col)) {
ai_clean <- ai_clean %>%
mutate(
sex_raw = as.character(.data[[sex_col]]),
sex = case_when(
sex_raw %in% c("1", "Male", "male", "M") ~ "Male",
sex_raw %in% c("2", "Female", "female", "F") ~ "Female",
TRUE ~ NA_character_
)
)
  ai_clean <- ai_clean %>%
    mutate(
      sex_raw = as.character(.data[[sex_col]]),
      sex = case_when(
        sex_raw %in% c("1", 1, "Male", "male", "M") ~ "Male",
        sex_raw %in% c("2", 2, "Female", "female", "F") ~ "Female",
        TRUE ~ NA_character_
      )
    )
}

# Detect benefit-related columns

benb_cols <- names(ai_clean)[str_detect(names(ai_clean), regex("^benb|benefit", ignore_case = TRUE))]
# glimpse(select(ai_clean, any_of(c("age", "sex", benb_cols))))

```

[*Inspect variables and "missingness"*]{.underline}

Before getting deeper into the data, we wanted to check for missing
variable types to understand parts of the questionnaire that were left
unanswered by parts of the survey population. This would give insight
into variables that are interesting, but that might have missing context
because of their lack of completion by survey respondents.

```{r}
name_map <- c(
  "governance_9" = "AI Governance – Who should ensure AI is used safely?",
  "benb_loan_7"  = "AI & Loans (Benefits) – How can AI loan risk prediction be beneficial?",
  "benb_rvc_4"   = "Robotic Vacuums (Benefits) – How are robotic vacuums beneficial?",
  "conb_loan_12" = "AI & Loans (Concerns) – What are your concerns about AI loan risk prediction?",
  "benb_aw_7"    = "Autonomous Weapons (Benefits) – How could autonomous weapons be beneficial?",
  "benb_ps_6"    = "Face Recognition in Policing (Benefits) – How could it benefit policing/surveillance?",
  "benb_bc_7"    = "Face Recognition at Borders (Benefits) – How could it benefit border control?",
  "benb_ccr_5"   = "AI & Climate Research (Benefits) – How could AI simulations help climate research?",
  "benb_can_6"   = "AI & Cancer Prediction (Benefits) – How could AI cancer prediction be beneficial?",
  "benb_smc_4"   = "Targeted Ads (Benefits) – How could targeted social media ads be beneficial?"
)
ai_missing <- tibble( variable = names(ai_clean), missing = map_int(ai_clean, ~ sum(is.na(.))), missing_pct = round(map_dbl(ai_clean, ~ mean(is.na(.))) * 100, 2) ) %>% arrange(desc(missing_pct)) %>% slice_head(n = 10)

ai_missing_disp <- ai_missing %>%
  mutate(
    variable_display = dplyr::recode(variable, !!!name_map, .default = variable)
    # If you want to keep the code too, use:
    # variable_display = dplyr::recode(variable, !!!name_map, .default = variable) |>
    #                    paste0(" (", variable, ")")
  ) %>%
  select(variable = variable_display, missing, missing_pct)

knitr::kable(ai_missing_disp, caption = "Top 10 Variables with the Highest Missingness")

# Detect BENB-like columns and coerce to 0/1 integers where possible
benb_cols <- names(ai_clean)[str_detect(names(ai_clean), regex("^benb|benb_job|benefit|benefits|benb_", ignore_case = TRUE))]
if (length(benb_cols) == 0) {
  benb_cols <- names(ai_clean)[str_detect(names(ai_clean), regex("benefit|benb|job", ignore_case = TRUE))]
}

if (length(benb_cols) > 0) {
  ai_clean <- ai_clean %>%
    mutate(across(all_of(benb_cols), ~ {
      x <- .
      # character encodings -> 1/0
      if (is.character(x)) {
        xl <- str_to_lower(str_trim(x))
        as.integer(if_else(xl %in% c("yes","y","1","selected","true","t"), 1L,
                   if_else(xl %in% c("no","n","0","false","f"), 0L, NA_integer_)))
      } else if (is.logical(x)) {
        as.integer(x)
      } else {
        # numeric-like: keep 1/0, else NA
        xi <- suppressWarnings(as.integer(x))
        if_else(is.na(xi), NA_integer_, xi)
      }
    }))
}


```

## Summary statistics

```{r summary-stats}
# Age summaries
if ("age" %in% names(ai_clean)) {
ai_clean %>%
summarise(
mean_age = mean(age, na.rm = TRUE),
median_age = median(age, na.rm = TRUE),
sd_age = sd(age, na.rm = TRUE),
min_age = min(age, na.rm = TRUE),
max_age = max(age, na.rm = TRUE)
)
  age_stats <- ai_clean %>%
    summarise(
      n = sum(!is.na(age)),
      mean_age = mean(age, na.rm = TRUE),
      median_age = median(age, na.rm = TRUE),
      sd_age = sd(age, na.rm = TRUE),
      min_age = min(age, na.rm = TRUE),
      max_age = max(age, na.rm = TRUE)
    )
  print(age_stats)
}

# Sex / gender distribution
if ("sex" %in% names(ai_clean)) {
ai_clean %>%
count(sex, sort = TRUE) %>%
mutate(percent = scales::percent(n / sum(n)))
  print(ai_clean %>% count(sex, sort = TRUE) %>% mutate(p = n / sum(n)))
}
```

## Visualizations and findings

[**Age Histogram**]{.underline}

```{r plots, fig.width=7, fig.height=4}
# Age histogram
if ("age" %in% names(ai_clean)) {
ggplot(filter(ai_clean, !is.na(age)), aes(age)) +
geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
labs(title = "Age Distribution", x = "Age", y = "Count")
  ggplot(filter(ai_clean, !is.na(age)), aes(age)) +
    geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
    labs(title = "Age distribution", x = "Age", y = "Count")
}

```

Most participants fall between their 20s and 50s, showing a good mix of
working-age adults. Fewer people over 70 took the survey, but the data
still includes a wide range of ages.

[**Gender Distribution**]{.underline}

```{r}
# Gender bar chart

# Sex / gender bar
if ("sex" %in% names(ai_clean)) {
ggplot(filter(ai_clean, !is.na(sex)), aes(x = fct_infreq(sex))) +
geom_bar(fill = "darkorange") +
coord_flip() +
labs(title = "Gender Distribution", x = NULL, y = "Count")
  ggplot(filter(ai_clean, !is.na(sex)), aes(x = fct_infreq(sex))) +
    geom_bar(fill = "darkorange") +
    coord_flip() +
    labs(title = "Sex / Gender distribution", x = NULL, y = "Count")
}
```

The chart shows that survey responses were evenly split between male and
female participants, giving a balanced view of public attitudes toward
AI. This balance helps ensure that later comparisons by gender aren’t
biased by unequal sample sizes.

[*Top 8 Perceived Benefits of AI*]{.underline}

**Derive Overall Benefit Tallies (Unweighted)**

```{r, eval = FALSE}
## Derive Overall Benefit Tallies (Unweighted)
# define which columns are your 0/1 benefit flags
benefit_cols <- grep("^benb_", names(ai_clean), value = TRUE)
# benefit_cols

df_long <- ai_clean %>%
  mutate(row_id = dplyr::row_number()) %>%
  tidyr::pivot_longer(
    cols = dplyr::all_of(benefit_cols),
    names_to   = "benefit_raw",
    values_to  = "selected",
    values_transform = list(selected = as.integer),  # coerce if needed
    values_drop_na   = TRUE
  ) %>%
  dplyr::filter(selected == 1L) %>%
  dplyr::mutate(benefit = benefit_raw) %>%
  dplyr::count(benefit) %>%
  dplyr::mutate(pct = n / sum(n))
```

```{r, eval= FALSE}
 # ai_clean %>% 
 #  dplyr::summarise(across(everything(), ~ all(. %in% c(0,1), na.rm = TRUE))) %>%
 # t() %>%
  # tibble::as_tibble(rownames = "var") %>%
  # dplyr::filter(value == TRUE)

```

```{r derive-overall-ben}
overall_ben <- ai_clean %>%
  tidyr::pivot_longer(all_of(benb_cols), names_to = "benefit", values_to = "selected",
                      values_transform = list(selected = as.integer),
                      values_drop_na = TRUE) %>%
  dplyr::filter(selected == 1L) %>%
  dplyr::count(benefit, sort = TRUE) %>%
  dplyr::mutate(pct = n / sum(n))
```

## 

```{r}
overall_ben <- overall_ben %>%
  mutate(
    benefit = str_to_lower(str_trim(benefit)),  # clean names first
    benefit_label = recode(benefit,
      "benb_mp01" = "Faster to unlock devices",
      "benb_mp02" = "Personal information more safe and secure",
      "benb_rvc03" = "Vacuuming possible for people who have trouble with manual tasks",
      "benb_can01" = "Earlier detection of cancer",
      "benb_ps01"  = "Identify wanted criminals and missing persons",
      "benb_vass02" = "People with difficulty using devices have better access to features",
      "benb_bc01"  = "Processing people at border control will be faster",
      "benb_rvc01" = "Tech can do vacuuming and save people time",
      .default = benefit  # keeps unmapped names so nothing breaks
    )
  )

```

```{r}
# Select Top 8 by count
top8 <- overall_ben %>%
  arrange(desc(n)) %>%
  slice_head(n = 8) %>%
  mutate(
    benefit_label = factor(
      benefit_label,
      levels = benefit_label[order(n, decreasing = TRUE)]
    )
  )

# Plot Top 8 with recoded labels
ggplot(top8, aes(x = n, y = benefit_label)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = n), hjust = -0.1, size = 3.7) +
  coord_cartesian(xlim = c(0, max(top8$n) * 1.15)) +
  labs(
    title = "Top 8 Benefits Selected",
    subtitle = "Raw number of respondents selecting each benefit",
    x = "Number of Respondents",
    y = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    axis.text.y = element_text(size = 10)
  )

```

The raw response data suggests that there is no single dominant theme in
how the public perceives the benefits of AI. Instead, respondents
highlight a wide range of advantages that span everyday life (e.g.,
vacuuming or unlocking devices), improved accessibility for people with
disabilities, enhanced security of personal information, and support for
public safety and border processes.

Overall, people seem to view AI as a set of practical tools that can
make daily life a bit easier, safer, and more efficient. But, their
perceived benefits are not focused on one area.

## Looking deeper - AI Usage for Accessibility Purposes

Because the perceived benefits of AI span a variety of domains, we next
focused our analysis on a specific application to better understand how
attitudes differ across demographic groups. We selected Robo-vacuums and
assistants as a case example: an everyday form of AI that can
meaningfully improve accessibility, particularly for individuals with
mobility limitations or other disabilities. Prior research has noted
that assistive cleaning technologies can reduce physical strain and
support independent living among vulnerable populations *([Cleaning Is
Caring,
n.d.)](https://www.cleaningiscaring.org/wpd/disability-friendly-cleaning/)*.
By examining responses related to Robo-vacuum and robo-assistants , we
explore whether age, gender, and other demographic characteristics
influence who recognizes these accessibility-related benefits of AI.

```{r}
# Identify column name automatically if needed
aware_var <- names(ai_clean)[str_detect(names(ai_clean), regex("aware_rca", ignore_case = TRUE))]
stopifnot(length(aware_var) == 1)

# Create age group bands if numeric age exists
age_group_df <- ai_clean %>%
  filter(!is.na(.data[[aware_var]])) %>%
  mutate(
    awareness = case_when(
      str_to_lower(str_trim(.data[[aware_var]])) %in% c("1","yes","y","true") ~ "Yes",
      str_to_lower(str_trim(.data[[aware_var]])) %in% c("0","no","n","false") ~ "No",
      TRUE ~ "Not sure"
    ),
    age_group = case_when(
      age < 25 ~ "18–24",
      age >= 25 & age <= 34 ~ "25–34",
      age >= 35 & age <= 44 ~ "35–44",
      age >= 45 & age <= 54 ~ "45–54",
      age >= 55 & age <= 64 ~ "55–64",
      age >= 65 ~ "65+",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(age_group))

aware_age <- ai_clean %>%
  select(age_group = age, aware_rca = aware_rca) %>%
  filter(!is.na(age_group), aware_rca %in% c(1,2,3)) %>%
  mutate(
    aware = case_when(
      aware_rca == 1 ~ 1L,        # Yes
      aware_rca %in% c(2,3) ~ 0L  # No + Not sure
    )
  ) %>%
  group_by(age_group) %>%
  summarise(pct_yes = mean(aware), .groups = "drop")

aware_age <- aware_age %>%
  mutate(age_num = as.numeric(age_group))  # Convert ordered factor to numeric for regression

ggplot(aware_age, aes(x = age_group, y = pct_yes)) +
  geom_col(fill = "#4B8EA8") +
   # Regression line
  geom_smooth(aes(x = age_num, y = pct_yes), 
              method = "lm", 
              se = FALSE, 
              color = "red", 
              size = 1,
              linetype = "dashed",
              inherit.aes = FALSE) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "Awareness of Robotic Care Assistants by Age",
    subtitle = "Percent who have heard of robotic care assistants",
    x = "Age Group", y = "% Aware"
  ) +
  theme_minimal(base_size = 14)
```

Although the bar heights suggest that awareness of robotic care
assistants increases with age, the regression line reveals a subtle
downward trend across older age groups. This discrepancy likely reflects
sample imbalance, as only 7.5% of respondents were 75+, resulting in
greater variability and reduced confidence when estimating awareness
levels for the oldest segment of the population.

```{r}
ai_clean %>%
  summarise(
    total = n(),
    n_75plus = sum(age >= 75, na.rm = TRUE),
    pct_75plus = round(n_75plus / total * 100, 1)
  )

```

```{r}
# Recode response values to labels
ai_plot <- ai_clean %>%
  select(age, bena_rca) %>%
  mutate(
    bena_rca = recode(as.character(bena_rca),
      "1" = "Very beneficial",
      "2" = "Fairly beneficial",
      "3" = "Not very beneficial",
      "4" = "Not at all beneficial",
      "5" = "Don't know",
      .default = NA_character_
    ),
    age_band = cut(
      age,
      breaks = c(17, 24, 34, 44, 54, 64, 74, Inf),
      labels = c("18–24", "25–34", "35–44", "45–54", "55–64", "65–74", "75+"),
      right = TRUE
    )
  ) %>%
  filter(!is.na(age_band), !is.na(bena_rca))

# Make simple bar chart
ggplot(ai_plot, aes(x = age_band, fill = bena_rca)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_brewer(palette = "Blues", direction = 1) +
  labs(
    title = "Perceived Benefit of Robotic Care Assistants by Age Group",
    subtitle = "Question: To what extent do you think the use of robotic care assistants will be beneficial?",
    x = "Age group",
    y = "Percent of respondents",
    fill = "Response"
  ) +
  theme_minimal(base_size = 12)
```

```{r}
concern_age <- ai_clean %>%
  select(age_group = age, concern = cona_rca) %>% 
  filter(!is.na(age_group) & concern %in% c(1,2,3,4,5)) %>%
  mutate(
    concerned = case_when(
      concern %in% c(1,2) ~ 1L,   # Very + Somewhat concerned
      concern %in% c(3,4,5) ~ 0L  # Not concerned + Don't know
    )
  ) %>%
  group_by(age_group) %>%
  summarise(
    pct_concern = mean(concerned, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

ggplot(concern_age, aes(x = age_group, y = pct_concern, group = 1)) +
  
  geom_col(fill = "#D95F02", alpha = 0.7) +

  # Regression line
  geom_smooth(aes(x = as.numeric(age_group), y = pct_concern),
              method = "lm", se = FALSE,
              color = "black", size = 1,
              linetype = "dashed",
              inherit.aes = FALSE) +
  
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0,1)) +
  
  labs(
    title = "Concern About Robotic Care Assistants by Age",
    subtitle = "Share of respondents who are very or somewhat concerned",
    x = "Age Group",
    y = "% Concerned"
  ) +

  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    panel.grid.minor = element_blank()
  )

```

Concern about robotic care assistants increases with age, indicating
that older respondents are especially wary of this technology. However,
elevated concern is evident across all age groups, showing that unease
toward robotic care assistants is a broadly shared sentiment and not
exclusively an older-adult issue.

```{r}
# Select all BENB_DC columns
benb_dc_data <- ai_clean %>%
  select(starts_with("benb_dc"))

# Convert to long format
benb_dc_long <- benb_dc_data %>%
  mutate(id = row_number()) %>%  # respondent ID
  pivot_longer(
    cols = starts_with("benb_dc"),
    names_to = "option",
    values_to = "selected"
  )

# Summarise counts and percentages
benb_dc_summary <- benb_dc_long %>%
  filter(!is.na(selected)) %>%
  group_by(option) %>%
  summarise(
    count = sum(selected == 1, na.rm = TRUE),
    total_respondents = n_distinct(id),
    percent = (count / total_respondents) * 100
  ) %>%
  ungroup() %>%
  # Clean up inconsistent naming
  mutate(
    option = tolower(option),
    option = str_replace_all(option, "benb[_]?dc[_]?", "benb_dc"),
    option = str_replace_all(option, "0+", "")
  ) %>%
  # Remove unwanted items
  filter(!option %in% c("benb_dc10", "benb_dc11")) %>%
  arrange(desc(percent))

# ---- PLOT ----
ggplot(benb_dc_summary, aes(x = reorder(option, percent), y = percent)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  scale_x_discrete(
    labels = c(
      "benb_dc1" = "Easier car travel",
      "benb_dc2" = "More time for other activities",
      "benb_dc3" = "More accurate driving",
      "benb_dc4" = "Fewer accidents",
      "benb_dc5" = "Easier for disabled or older drivers",
      "benb_dc6" = "Saves money on drivers",
      "benb_dc7" = "Something else",
      "benb_dc8" = "None of these",
      "benb_dc9" = "Don't know"
    )
  ) +
  labs(
    title = "Perceived Benefits of Driverless Cars",
    x = NULL,
    y = "Percentage of Respondents (%)"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10),
    plot.title = element_text(face = "bold", size = 14)
  )
```

## Exploring benefit score vs age (bridge to modeling)

```{r}
if (length(benb_cols) > 0 && "age" %in% names(ai_clean)) {

  ai_model_df <- ai_clean %>%
    mutate(
      benefit_score = rowSums(across(all_of(benb_cols)), na.rm = TRUE)
    ) %>%
    select(age, benefit_score) %>%
    filter(!is.na(age), !is.na(benefit_score))

  ggplot(ai_model_df, aes(x = age, y = benefit_score)) +
    geom_point(alpha = 0.2) +
    geom_smooth(method = "lm", se = FALSE, color = "red") +
    labs(
      title = "Total Number of Benefits Selected vs Age",
      subtitle = "A simple illustration of how a regression model may treat benefit score as an outcome",
      x = "Age",
      y = "Benefit score (count of benefits selected)"
    ) +
    theme_minimal()
}
```
The scatterplot shows that although people across all ages select a wide range of AI benefits, there is no strong linear relationship between age and the total number of benefits selected. The regression line is almost flat, suggesting that older adults are not necessarily more or less likely than younger adults to see AI as beneficial overall.
However, the large spread in benefit scores at every age indicates that age alone cannot explain variation in attitudes. This reinforces why a regression model is needed: to test how age interacts with other factors such as gender, education, socio-economic status, and awareness—to more accurately predict who perceives AI as beneficial.

```{r}
## Prepare data for regression

# 1. Overall benefit score: sum across all benb_ columns
ai_clean$model_benefit_score <- rowSums(ai_clean[, benb_cols], na.rm = TRUE)

# 2. Recode outcomes for Robotic Care Assistants (RCA)
ai_clean <- ai_clean %>%
  mutate(
    rca_beneficial = case_when(
      bena_rca %in% c(1, 2) ~ 1,   # Very or fairly beneficial
      bena_rca %in% c(3, 4) ~ 0,   # Not very / not at all beneficial
      TRUE ~ NA_real_
    ),
    rca_concerned = case_when(
      cona_rca %in% c(1, 2) ~ 1,   # Very or somewhat concerned
      cona_rca %in% c(3, 4, 5) ~ 0, # Not very / not at all / don't know
      TRUE ~ NA_real_
    )
  )

# 3. Build modeling dataset with the correct column names
model_df <- ai_clean %>%
  select(
    age,
    sex,
    rs_education_level,  # <-- this is the real name in your data
    dr_socio_eco,        # <-- socio-economic group
    model_benefit_score,
    rca_beneficial,
    rca_concerned
  ) %>%
  filter(!is.na(age), !is.na(sex))

head(model_df)


```
```{r}
# Convert categorical predictors to factors
model_df <- model_df %>%
  mutate(
    sex = factor(sex),
    rs_education_level = factor(rs_education_level,
                                labels = c("Degree", "Non-degree", "No qualifications")),
    dr_socio_eco = factor(dr_socio_eco)
  )
```

## Model 1: Benefit Score ~ Age

```{r}
model1 <- lm(model_benefit_score ~ age, data = model_df)
summary(model1)
```

## Model 2: Logistic Regression (Who thinks RCAs are beneficial?)

```{r}
model2 <- glm(
  rca_beneficial ~ age + sex + rs_education_level + dr_socio_eco,
  data = model_df,
  family = binomial()
)
summary(model2)
```
## Model 2: graph

```{r coef_plot_model2, fig.width=7, fig.height=5}
m2_tidy <- broom::tidy(model2, conf.int = TRUE, exponentiate = TRUE) %>%
  dplyr::filter(term != "(Intercept)") %>%
  dplyr::mutate(
    term = dplyr::recode(term,
                         "sexMale" = "Male (vs Female)",
                         "age" = "Age (per year)",
                         "rs_education_levelNon-degree" = "Non-degree (vs Degree)",
                         "rs_education_levelNo qualifications" = "No qualifications (vs Degree)",
                         "dr_socio_eco2" = "SEC 2",
                         "dr_socio_eco3" = "SEC 3",
                         "dr_socio_eco4" = "SEC 4",
                         "dr_socio_eco5" = "SEC 5",
                         "dr_socio_eco6" = "SEC 6",
                         "dr_socio_eco7" = "SEC 7"
                        ),
    term = forcats::fct_reorder(term, estimate)
  )

ggplot(m2_tidy, aes(x = term, y = estimate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.1) +
  coord_flip() +
  labs(
    title = "Odds Ratios for Viewing RCAs as Beneficial",
    x = NULL,
    y = "Odds Ratio (95% CI)"
  ) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  theme_minimal()
```

## Model 3: Logistic Regression (Who is concerned about RCAs?)

```{r}
model3 <- glm(
  rca_concerned ~ age + sex + rs_education_level + dr_socio_eco,
  data = model_df,
  family = binomial()
)
summary(model3)
```

## Graph

```{r coef_plot_model3, fig.width=7, fig.height=5}
m3_tidy <- broom::tidy(model3, conf.int = TRUE, exponentiate = TRUE) %>%
  dplyr::filter(term != "(Intercept)") %>%
  dplyr::mutate(
    term = dplyr::recode(
      term,
      "sexMale" = "Male (vs Female)",
      "age" = "Age (per year)",
      "rs_education_levelNon-degree" = "Non-degree (vs Degree)",
      "rs_education_levelNo qualifications" = "No qualifications (vs Degree)",
      "dr_socio_eco2" = "SEC 2",
      "dr_socio_eco3" = "SEC 3",
      "dr_socio_eco4" = "SEC 4",
      "dr_socio_eco5" = "SEC 5",
      "dr_socio_eco6" = "SEC 6",
      "dr_socio_eco7" = "SEC 7"
    ),
    term = forcats::fct_reorder(term, estimate)
  )

ggplot(m3_tidy, aes(x = term, y = estimate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.1) +
  coord_flip() +
  labs(
    title = "Odds Ratios for Concern About RCAs",
    x = NULL,
    y = "Odds Ratio (95% CI)"
  ) +
  geom_hline(yintercept = 1, linetype = "dashed") +
  theme_minimal()
```

## Regression Results 


### Model 1: Benefit Score ~ Age
Age is not a significant predictor of the total number of AI benefits a respondent selects. The slope is very close to zero, and the model explains almost none of the variation in benefit scores. This confirms our earlier scatterplot: people across all ages report similar levels of perceived benefit.

### Model 2: RCA Beneficial ~ Age + Sex + Education + SEC
Age, gender, and education significantly predict whether respondents view robotic care assistants as beneficial. Older adults are less likely to view RCAs positively, while men are more likely to see benefits. Respondents with lower education levels and some lower socio-economic groups also show reduced likelihood of perceiving RCAs as beneficial. These results suggest that demographic differences meaningfully shape attitudes toward assistive robotic technologies

### Model 3: RCA Concerned ~ Age + Sex + Education + SEC
Age and gender significantly predict concern about robotic care assistants. Older adults are more likely to be concerned, while men are substantially less likely to express concern. Education and most socio-economic categories do not show significant effects. Concern about RCAs appears to be driven mainly by age and gender differences.

## Conclusion 

Overall, our analysis shows that while artificial intelligence is now woven into everyday life, people vary in how they perceive its usefulness—especially when AI is applied to accessibility-related tasks. Across examples such as robotic vacuum cleaners, robotic care assistants, and driverless cars, many respondents recognize practical benefits: saving time, reducing physical strain, and supporting independent living for older adults or individuals with mobility limitations. These descriptive patterns reflect a general optimism toward the potential of AI to make daily activities easier and more efficient.
However, the regression results reveal that these attitudes are not evenly distributed across demographic groups. Age and gender consistently emerged as significant predictors. Older adults were less likely to view robotic care assistants as beneficial and more likely to express concern, suggesting heightened caution among populations who could arguably benefit most from assistive technologies. Gender gaps were also notable—men had higher odds of seeing RCAs as beneficial and significantly lower odds of being concerned compared to women. While education and socio-economic status showed some influence, their effects were smaller and less consistent.
Together, these findings suggest that public attitudes toward AI accessibility tools are shaped not only by the technology itself but also by the demographic and social experiences people bring to it. As AI systems become more embedded in health, caregiving, and household activities, understanding these differences is essential. Designing inclusive technologies will require paying attention not only to functional performance but also to the social and emotional factors—such as trust, familiarity, and perceived risk—that shape how different groups evaluate AI in everyday life

## Future Direction 

The models in this notebook focus on broad demographic predictors of how people
view robotic care assistants. A natural next step would be to incorporate more
direct measures of accessibility and disability (for example, disability status
and caring responsibilities) to see whether those who might benefit most from
assistive technologies are more likely to view them as useful or concerning. It
would also be valuable to add measures of technology experience and confidence,
and to explore interaction terms (such as age × disability or gender × tech
comfort) to test whether demographic groups respond differently depending on
their everyday relationship with technology. Finally, future work could apply
survey weights and more flexible models (e.g., logistic regression with
interactions or multilevel models) to better generalize these findings to the
UK population.
