---
editor_options: 
  markdown: 
    wrap: 72
---

\@\@ -5,114 +5,153 \@\@ date: "`r Sys.Date()`" output: html_document ---

```{r setup, include = FALSE}
# Load all required packages
library(ggplot2)
library(dplyr)
library(forcats)
library(tidyr)
library(stringr)
library(scales)
library(readr)
library(janitor)
library(corrplot)

# Global knit options
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
set.seed(2025)

```

```{r, include=FALSE}
library(scales)
# setup: load libraries and options
library(tidyverse)
library(readr)
library(haven)
library(janitor)
library(scales)
library(stringr)
library(dplyr)
library(tidyr)


knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
set.seed(2025)
```

## Introduction

**Background**

In recent years, artificial intelligence (AI) has shifted from a source
of uncertainty and risk to a common feature of everyday life. From
writing assistants and chat bots to algorithmic recommendations in
education, finance, and entertainment, AI has become deeply embedded in
social and professional contexts. Understanding how people perceive its
benefits and risks is crucial for policymakers and organizations
designing AI systems that are equitable and trustworthy. In recent
years, artificial intelligence (AI) has undergone a profound shift in
public perception. Once seen primarily as a source of uncertainty or
even risk, AI is now becoming something people feel increasingly
comfortable with—if not socially or professionally expected to use. From
writing assistants and chatbots to algorithmic recommendations in
education, finance, and entertainment, AI has become an embedded feature
of everyday life.

**Dataset and preprocessing**

To ground our analysis, we use the Public Attitudes to AI (PAI) data set
produced by the Ada Lovelace Institute (UK). This nationally
representative survey of approximately 4,000 adults captures attitudes
toward AI across multiple domains—such as hiring, healthcare, and
policing. To ground our analysis, we use the Public Attitudes to AI
dataset produced by the Ada Lovelace Institute (UK). This nationally
representative survey of \~4,000 adults provides perceptions across
multiple AI application domains. Our aim here is a reproducible initial
exploration: inspect structure and missingness, perform light cleaning,
compute simple summaries, and visualise perceived benefits of AI in
hiring.

This project uses the preprocessed version of the data set provided by
the Ada Lovelace Institute. Our group utilized similar cleaning
techniques based on the descriptive analysis code by the Ada Lovelace
Institute.

**Goal:** Our project aims to replicate the results of the Ada Lovelace
Institution, but to also consider gender, age, and other demographics
that impacted survey respondents' perceived benefits of AI.

## Explaining the Dataset

**Loading raw data set**

```{r}
ai_path <- "how-do-people-feel-about-ai-main/262400371_Public Attitudes to AI Raw Data NA (1).csv"
if (!file.exists(ai_path)) stop("Data file not found: ", ai_path)
ai_survey1 <- read_csv(ai_path, show_col_types = FALSE)
head(ai_survey1, 10)
```

**Understanding the variables and tidying up**

The dataset includes demographic variables such as age, sex, and region
along with multiple binary columns (beginning with 'benb\_')
representing whether respondents selected each perceived benefit. Below
we summarize the key variables used in our analysis

```{r}
tibble::tribble(
~variable,       ~description,                                    ~type,
"age",           "Respondent age in years",                       "Numeric",
"sex",           "Self-identified gender (Male/Female/Other)",    "Categorical",
"benb_*",        "Binary indicators for perceived AI benefits",   "Integer (0/1)",
"weight",        "Survey weight (if applicable)",                 "Numeric"
) %>%
knitr::kable(caption = "Key variables in the preprocessed dataset")
```

[*Recoding technology labels*]{.underline}

```{r}
recode_technologies <- function(Technology_specific) {
  recode(Technology_specific,
                  "WEL" = "Welfare eligibility",
                "JOB" =  "Job eligibility",
                "CAN" =  "Cancer risk",
                "LOAN" = "Loan repayment risk", 
                "MP" =  "Mobile phone unlock",
                 "BC" = "Border control",
                 "PS" = "Police surveillance",

                "AW"  = "Autonomous weapons", 
                "DC" =  "Driverless cars",
                "RCA" = "Robotic care assistant",
               "RVC" ="Robotic vacuum cleaner",
                                                                         
               "SMC"=  "Targeted consumer ads",
                "SMP" = "Targeted political ads",
                                                                         
                "VASS"=  "Smart speaker",
                "VAH"=  "Virtual healthcare assistant",
                                 
              "CCR"=   "Climate change simulation",
                "VR"=  "Education simulations")
  
}
```

[*List of demographic variables of interest*]{.underline}

```{r}
demog <- ai_survey1 %>%
  haven::zap_labels() %>%
  dplyr::select(
  Respondent_ID, PV25_Weight,
  RS_Ethnicity_Reduced,
  CS_Age_Band,
  RS_Sex,
  RS_EducationLevel,
  DR_Socio_Eco
  ) %>%
  mutate(RS_Ethnicity_Combined = 
           dplyr::recode(RS_Ethnicity_Reduced,
             "1" = "White",
             "2" = "Asian or Asian British",
             "3" = "Black British, Caribbean or African",
             "4" = "Any other ethnic group"
           )) %>%
  mutate(CS_Age_Band = 
           dplyr::recode(CS_Age_Band,
             "1" = "18-24 yrs",
             "2" = "25-34 yrs",
             "3" = "35-44 yrs",
             "4" = "45-54 yrs",
             "5" = "55-64 yrs",
             "6" = "65-74 yrs",
             "7" = "75+ yrs"
           )) %>%
  mutate(RS_Sex = 
           dplyr::recode(RS_Sex,
             "1" = "Male",
             "2" = "Female")) %>%
  mutate(RS_EducationLevel = 
           dplyr::recode(RS_EducationLevel,
             "1" = "Degree level qualification(s)",
             "2" = "Non-degree level qualifications",
             "3" = "No academic or vocational qualifications")) %>%
  mutate(DR_Socio_Eco = 
           dplyr::recode(DR_Socio_Eco,
             "1" = "SEC1, 2",
             "2" = "SEC3",
             "3" = "SEC4",
             "4" = "SEC5",
             "5" = "SEC6, 7",
             "6" = "SEC8",
             "7" = "Students")) %>%
  pivot_longer(names_to = "Q_Demo",
               values_to = "A_Demo",
               RS_Ethnicity_Combined:DR_Socio_Eco)

# list of levels 
demo_list <- c("White",
  "Asian or Asian British",
  "Black British, Caribbean or African", 
  "Any other ethnic group", 
  
  "18-24 yrs",
  "25-34 yrs",
  "35-44 yrs",
  "45-54 yrs",
  "55-64 yrs",
  "65-74 yrs",
  "75+ yrs", 
  
  "Male",
  "Female",
  "Identify in another way",
  
  "Degree level qualification(s)",
  "Non-degree level qualifications",
  "No academic or vocational qualifications",
  
  "SEC1, 2",
  "SEC3",
  "SEC4",
  "SEC5",
  "SEC6, 7",
  "SEC8",
  "Students", 
  
  "Total" )

if ("age" %in% names(ai_survey1)) {
ggplot(filter(ai, !is.na(age)), aes(age)) +
geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
labs(title = "Age Distribution of Respondents", x = "Age (years)", y = "Count")
}

```

[*Identify and prepare key column key and creating "ai_clean"
dataset*]{.underline}

We standardize variable names, detect age and gender columns and
identify all AI benefit-related columns (benb\_).

```{r}
## Basic cleaning and derived columns
ai <- read_csv(ai_path, show_col_types = FALSE)
# Standardise names and add row id
ai_clean <- ai_survey1 %>%
  clean_names() %>%
  mutate(row_id = row_number())

ai_clean <- ai %>%
clean_names() %>%
mutate(row_id = row_number())

# Detect age column

age_col <- names(ai_clean)[str_detect(names(ai_clean), regex("age", ignore_case = TRUE))][1]
# Detect age column (common names cs_age, csage, cs_age_yrs, age)
age_col <- names(ai_clean)[str_detect(names(ai_clean), regex("\\b(cs_age|csage|age|age_years|cs_age_yrs)\\b", ignore_case = TRUE))] %>% .[1]
if (!is.na(age_col)) {
ai_clean <- ai_clean %>%
mutate(age = as.integer(.data[[age_col]])) %>%
filter(is.na(age) | age >= 17)
  ai_clean <- ai_clean %>%
    mutate(age = as.integer(.data[[age_col]])) %>%
    filter(is.na(age) | age >= 17)  # remove implausible 16 if present
}

# Detect sex/gender column

sex_col <- names(ai_clean)[str_detect(names(ai_clean), regex("sex|gender", ignore_case = TRUE))][1]
sex_col <- names(ai_clean)[str_detect(names(ai_clean), regex("^rs_sex$|sex|gender", ignore_case = TRUE))] %>% .[1]
if (!is.na(sex_col)) {
ai_clean <- ai_clean %>%
mutate(
sex_raw = as.character(.data[[sex_col]]),
sex = case_when(
sex_raw %in% c("1", "Male", "male", "M") ~ "Male",
sex_raw %in% c("2", "Female", "female", "F") ~ "Female",
TRUE ~ NA_character_
)
)
  ai_clean <- ai_clean %>%
    mutate(
      sex_raw = as.character(.data[[sex_col]]),
      sex = case_when(
        sex_raw %in% c("1", 1, "Male", "male", "M") ~ "Male",
        sex_raw %in% c("2", 2, "Female", "female", "F") ~ "Female",
        TRUE ~ NA_character_
      )
    )
}

# Detect benefit-related columns

benb_cols <- names(ai_clean)[str_detect(names(ai_clean), regex("^benb|benefit", ignore_case = TRUE))]
# glimpse(select(ai_clean, any_of(c("age", "sex", benb_cols))))

```

[*Inspect variables and "missingness"*]{.underline}

Before getting deeper into the data, we wanted to check for missing
variable types to understand parts of the questionnaire that were left
unanswered by parts of the survey population. This would give insight
into variables that are interesting, but that might have missing context
because of their lack of completion by survey respondents.

```{r}
name_map <- c(
  "governance_9" = "AI Governance – Who should ensure AI is used safely?",
  "benb_loan_7"  = "AI & Loans (Benefits) – How can AI loan risk prediction be beneficial?",
  "benb_rvc_4"   = "Robotic Vacuums (Benefits) – How are robotic vacuums beneficial?",
  "conb_loan_12" = "AI & Loans (Concerns) – What are your concerns about AI loan risk prediction?",
  "benb_aw_7"    = "Autonomous Weapons (Benefits) – How could autonomous weapons be beneficial?",
  "benb_ps_6"    = "Face Recognition in Policing (Benefits) – How could it benefit policing/surveillance?",
  "benb_bc_7"    = "Face Recognition at Borders (Benefits) – How could it benefit border control?",
  "benb_ccr_5"   = "AI & Climate Research (Benefits) – How could AI simulations help climate research?",
  "benb_can_6"   = "AI & Cancer Prediction (Benefits) – How could AI cancer prediction be beneficial?",
  "benb_smc_4"   = "Targeted Ads (Benefits) – How could targeted social media ads be beneficial?"
)
ai_missing <- tibble( variable = names(ai_clean), missing = map_int(ai_clean, ~ sum(is.na(.))), missing_pct = round(map_dbl(ai_clean, ~ mean(is.na(.))) * 100, 2) ) %>% arrange(desc(missing_pct)) %>% slice_head(n = 10)

ai_missing_disp <- ai_missing %>%
  mutate(
    variable_display = dplyr::recode(variable, !!!name_map, .default = variable)
    # If you want to keep the code too, use:
    # variable_display = dplyr::recode(variable, !!!name_map, .default = variable) |>
    #                    paste0(" (", variable, ")")
  ) %>%
  select(variable = variable_display, missing, missing_pct)

knitr::kable(ai_missing_disp, caption = "Top 10 Variables with the Highest Missingness")

# Detect BENB-like columns and coerce to 0/1 integers where possible
benb_cols <- names(ai_clean)[str_detect(names(ai_clean), regex("^benb|benb_job|benefit|benefits|benb_", ignore_case = TRUE))]
if (length(benb_cols) == 0) {
  benb_cols <- names(ai_clean)[str_detect(names(ai_clean), regex("benefit|benb|job", ignore_case = TRUE))]
}

if (length(benb_cols) > 0) {
  ai_clean <- ai_clean %>%
    mutate(across(all_of(benb_cols), ~ {
      x <- .
      # character encodings -> 1/0
      if (is.character(x)) {
        xl <- str_to_lower(str_trim(x))
        as.integer(if_else(xl %in% c("yes","y","1","selected","true","t"), 1L,
                   if_else(xl %in% c("no","n","0","false","f"), 0L, NA_integer_)))
      } else if (is.logical(x)) {
        as.integer(x)
      } else {
        # numeric-like: keep 1/0, else NA
        xi <- suppressWarnings(as.integer(x))
        if_else(is.na(xi), NA_integer_, xi)
      }
    }))
}


```

## Summary statistics

```{r summary-stats}
# Age summaries
if ("age" %in% names(ai_clean)) {
ai_clean %>%
summarise(
mean_age = mean(age, na.rm = TRUE),
median_age = median(age, na.rm = TRUE),
sd_age = sd(age, na.rm = TRUE),
min_age = min(age, na.rm = TRUE),
max_age = max(age, na.rm = TRUE)
)
  age_stats <- ai_clean %>%
    summarise(
      n = sum(!is.na(age)),
      mean_age = mean(age, na.rm = TRUE),
      median_age = median(age, na.rm = TRUE),
      sd_age = sd(age, na.rm = TRUE),
      min_age = min(age, na.rm = TRUE),
      max_age = max(age, na.rm = TRUE)
    )
  print(age_stats)
}

# Sex / gender distribution
if ("sex" %in% names(ai_clean)) {
ai_clean %>%
count(sex, sort = TRUE) %>%
mutate(percent = scales::percent(n / sum(n)))
  print(ai_clean %>% count(sex, sort = TRUE) %>% mutate(p = n / sum(n)))
}
```

## Visualizations and findings

[**Age Histogram**]{.underline}

```{r plots, fig.width=7, fig.height=4}
# Age histogram
if ("age" %in% names(ai_clean)) {
ggplot(filter(ai_clean, !is.na(age)), aes(age)) +
geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
labs(title = "Age Distribution", x = "Age", y = "Count")
  ggplot(filter(ai_clean, !is.na(age)), aes(age)) +
    geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +
    labs(title = "Age distribution", x = "Age", y = "Count")
}

```

Most participants fall between their 20s and 50s, showing a good mix of
working-age adults. Fewer people over 70 took the survey, but the data
still includes a wide range of ages.

[**Gender Distribution**]{.underline}

```{r}
# Gender bar chart

# Sex / gender bar
if ("sex" %in% names(ai_clean)) {
ggplot(filter(ai_clean, !is.na(sex)), aes(x = fct_infreq(sex))) +
geom_bar(fill = "darkorange") +
coord_flip() +
labs(title = "Gender Distribution", x = NULL, y = "Count")
  ggplot(filter(ai_clean, !is.na(sex)), aes(x = fct_infreq(sex))) +
    geom_bar(fill = "darkorange") +
    coord_flip() +
    labs(title = "Sex / Gender distribution", x = NULL, y = "Count")
}
```

The chart shows that survey responses were evenly split between male and
female participants, giving a balanced view of public attitudes toward
AI. This balance helps ensure that later comparisons by gender aren’t
biased by unequal sample sizes.

[*Top 8 Perceived Benefits of AI*]{.underline}

**Derive Overall Benefit Tallies (Unweighted)**

```{r, eval = FALSE}
## Derive Overall Benefit Tallies (Unweighted)
# define which columns are your 0/1 benefit flags
benefit_cols <- grep("^benb_", names(ai_clean), value = TRUE)
# benefit_cols

df_long <- ai_clean %>%
  mutate(row_id = dplyr::row_number()) %>%
  tidyr::pivot_longer(
    cols = dplyr::all_of(benefit_cols),
    names_to   = "benefit_raw",
    values_to  = "selected",
    values_transform = list(selected = as.integer),  # coerce if needed
    values_drop_na   = TRUE
  ) %>%
  dplyr::filter(selected == 1L) %>%
  dplyr::mutate(benefit = benefit_raw) %>%
  dplyr::count(benefit) %>%
  dplyr::mutate(pct = n / sum(n))
```

```{r, eval= FALSE}
 # ai_clean %>% 
 #  dplyr::summarise(across(everything(), ~ all(. %in% c(0,1), na.rm = TRUE))) %>%
 # t() %>%
  # tibble::as_tibble(rownames = "var") %>%
  # dplyr::filter(value == TRUE)

```

```{r derive-overall-ben}
overall_ben <- ai_clean %>%
  tidyr::pivot_longer(all_of(benb_cols), names_to = "benefit", values_to = "selected",
                      values_transform = list(selected = as.integer),
                      values_drop_na = TRUE) %>%
  dplyr::filter(selected == 1L) %>%
  dplyr::count(benefit, sort = TRUE) %>%
  dplyr::mutate(pct = n / sum(n))
```

## 

```{r}
overall_ben <- overall_ben %>%
  mutate(
    benefit = str_to_lower(str_trim(benefit)),  # clean names first
    benefit_label = recode(benefit,
      "benb_mp01" = "Faster to unlock devices",
      "benb_mp02" = "Personal information more safe and secure",
      "benb_rvc03" = "Vacuuming possible for people who have trouble with manual tasks",
      "benb_can01" = "Earlier detection of cancer",
      "benb_ps01"  = "Identify wanted criminals and missing persons",
      "benb_vass02" = "People with difficulty using devices have better access to features",
      "benb_bc01"  = "Processing people at border control will be faster",
      "benb_rvc01" = "Tech can do vacuuming and save people time",
      .default = benefit  # keeps unmapped names so nothing breaks
    )
  )

```

```{r}
# Select Top 8 by count
top8 <- overall_ben %>%
  arrange(desc(n)) %>%
  slice_head(n = 8) %>%
  mutate(
    benefit_label = factor(
      benefit_label,
      levels = benefit_label[order(n, decreasing = TRUE)]
    )
  )

# Plot Top 8 with recoded labels
ggplot(top8, aes(x = n, y = benefit_label)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = n), hjust = -0.1, size = 3.7) +
  coord_cartesian(xlim = c(0, max(top8$n) * 1.15)) +
  labs(
    title = "Top 8 Benefits Selected",
    subtitle = "Raw number of respondents selecting each benefit",
    x = "Number of Respondents",
    y = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    axis.text.y = element_text(size = 10)
  )

```

The raw response data suggests that there is no single dominant theme in
how the public perceives the benefits of AI. Instead, respondents
highlight a wide range of advantages that span everyday life (e.g.,
vacuuming or unlocking devices), improved accessibility for people with
disabilities, enhanced security of personal information, and support for
public safety and border processes.

Overall, people seem to view AI as a set of practical tools that can
make daily life a bit easier, safer, and more efficient. But, their
perceived benefits are not focused on one area.

## Looking deeper - AI Usage for Accessibility Purposes

Because the perceived benefits of AI span a variety of domains, we next
focused our analysis on a specific application to better understand how
attitudes differ across demographic groups. We selected Robo-vacuums and
assistants as a case example: an everyday form of AI that can
meaningfully improve accessibility, particularly for individuals with
mobility limitations or other disabilities. Prior research has noted
that assistive cleaning technologies can reduce physical strain and
support independent living among vulnerable populations *([Cleaning Is
Caring,
n.d.)](https://www.cleaningiscaring.org/wpd/disability-friendly-cleaning/)*.
By examining responses related to Robo-vacuum and robo-assistants , we
explore whether age, gender, and other demographic characteristics
influence who recognizes these accessibility-related benefits of AI.

```{r}
# Identify column name automatically if needed
aware_var <- names(ai_clean)[str_detect(names(ai_clean), regex("aware_rca", ignore_case = TRUE))]
stopifnot(length(aware_var) == 1)

# Create age group bands if numeric age exists
age_group_df <- ai_clean %>%
  filter(!is.na(.data[[aware_var]])) %>%
  mutate(
    awareness = case_when(
      str_to_lower(str_trim(.data[[aware_var]])) %in% c("1","yes","y","true") ~ "Yes",
      str_to_lower(str_trim(.data[[aware_var]])) %in% c("0","no","n","false") ~ "No",
      TRUE ~ "Not sure"
    ),
    age_group = case_when(
      age < 25 ~ "18–24",
      age >= 25 & age <= 34 ~ "25–34",
      age >= 35 & age <= 44 ~ "35–44",
      age >= 45 & age <= 54 ~ "45–54",
      age >= 55 & age <= 64 ~ "55–64",
      age >= 65 ~ "65+",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(age_group))

aware_age <- ai_clean %>%
  select(age_group = age, aware_rca = aware_rca) %>%
  filter(!is.na(age_group), aware_rca %in% c(1,2,3)) %>%
  mutate(
    aware = case_when(
      aware_rca == 1 ~ 1L,        # Yes
      aware_rca %in% c(2,3) ~ 0L  # No + Not sure
    )
  ) %>%
  group_by(age_group) %>%
  summarise(pct_yes = mean(aware), .groups = "drop")

aware_age <- aware_age %>%
  mutate(age_num = as.numeric(age_group))  # Convert ordered factor to numeric for regression

ggplot(aware_age, aes(x = age_group, y = pct_yes)) +
  geom_col(fill = "#4B8EA8") +
   # Regression line
  geom_smooth(aes(x = age_num, y = pct_yes), 
              method = "lm", 
              se = FALSE, 
              color = "red", 
              size = 1,
              linetype = "dashed",
              inherit.aes = FALSE) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "Awareness of Robotic Care Assistants by Age",
    subtitle = "Percent who have heard of robotic care assistants",
    x = "Age Group", y = "% Aware"
  ) +
  theme_minimal(base_size = 14)
```

Although the bar heights suggest that awareness of robotic care
assistants increases with age, the regression line reveals a subtle
downward trend across older age groups. This discrepancy likely reflects
sample imbalance, as only 7.5% of respondents were 75+, resulting in
greater variability and reduced confidence when estimating awareness
levels for the oldest segment of the population.

```{r}
ai_clean %>%
  summarise(
    total = n(),
    n_75plus = sum(age >= 75, na.rm = TRUE),
    pct_75plus = round(n_75plus / total * 100, 1)
  )

```

```{r}
# Recode response values to labels
ai_plot <- ai_clean %>%
  select(age, bena_rca) %>%
  mutate(
    bena_rca = recode(as.character(bena_rca),
      "1" = "Very beneficial",
      "2" = "Fairly beneficial",
      "3" = "Not very beneficial",
      "4" = "Not at all beneficial",
      "5" = "Don't know",
      .default = NA_character_
    ),
    age_band = cut(
      age,
      breaks = c(17, 24, 34, 44, 54, 64, 74, Inf),
      labels = c("18–24", "25–34", "35–44", "45–54", "55–64", "65–74", "75+"),
      right = TRUE
    )
  ) %>%
  filter(!is.na(age_band), !is.na(bena_rca))

# Make simple bar chart
ggplot(ai_plot, aes(x = age_band, fill = bena_rca)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_brewer(palette = "Blues", direction = 1) +
  labs(
    title = "Perceived Benefit of Robotic Care Assistants by Age Group",
    subtitle = "Question: To what extent do you think the use of robotic care assistants will be beneficial?",
    x = "Age group",
    y = "Percent of respondents",
    fill = "Response"
  ) +
  theme_minimal(base_size = 12)
```

```{r}
concern_age <- ai_clean %>%
  select(age_group = age, concern = cona_rca) %>% 
  filter(!is.na(age_group) & concern %in% c(1,2,3,4,5)) %>%
  mutate(
    concerned = case_when(
      concern %in% c(1,2) ~ 1L,   # Very + Somewhat concerned
      concern %in% c(3,4,5) ~ 0L  # Not concerned + Don't know
    )
  ) %>%
  group_by(age_group) %>%
  summarise(
    pct_concern = mean(concerned, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

ggplot(concern_age, aes(x = age_group, y = pct_concern, group = 1)) +
  
  geom_col(fill = "#D95F02", alpha = 0.7) +

  # Regression line
  geom_smooth(aes(x = as.numeric(age_group), y = pct_concern),
              method = "lm", se = FALSE,
              color = "black", size = 1,
              linetype = "dashed",
              inherit.aes = FALSE) +
  
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0,1)) +
  
  labs(
    title = "Concern About Robotic Care Assistants by Age",
    subtitle = "Share of respondents who are very or somewhat concerned",
    x = "Age Group",
    y = "% Concerned"
  ) +

  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    panel.grid.minor = element_blank()
  )

```

Concern about robotic care assistants increases with age, indicating
that older respondents are especially wary of this technology. However,
elevated concern is evident across all age groups, showing that unease
toward robotic care assistants is a broadly shared sentiment and not
exclusively an older-adult issue.

```{r}

```

```{r}
# Select all BENB_DC columns
benb_dc_data <- ai_clean %>%
  select(starts_with("benb_dc"))

# Convert to long format
benb_dc_long <- benb_dc_data %>%
  mutate(id = row_number()) %>%  # respondent ID
  pivot_longer(
    cols = starts_with("benb_dc"),
    names_to = "option",
    values_to = "selected"
  )

# Summarise counts and percentages
benb_dc_summary <- benb_dc_long %>%
  filter(!is.na(selected)) %>%
  group_by(option) %>%
  summarise(
    count = sum(selected == 1, na.rm = TRUE),
    total_respondents = n_distinct(id),
    percent = (count / total_respondents) * 100
  ) %>%
  ungroup() %>%
  # Clean up inconsistent naming
  mutate(
    option = tolower(option),
    option = str_replace_all(option, "benb[_]?dc[_]?", "benb_dc"),
    option = str_replace_all(option, "0+", "")
  ) %>%
  # Remove unwanted items
  filter(!option %in% c("benb_dc10", "benb_dc11")) %>%
  arrange(desc(percent))

# ---- PLOT ----
ggplot(benb_dc_summary, aes(x = reorder(option, percent), y = percent)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  scale_x_discrete(
    labels = c(
      "benb_dc1" = "Easier car travel",
      "benb_dc2" = "More time for other activities",
      "benb_dc3" = "More accurate driving",
      "benb_dc4" = "Fewer accidents",
      "benb_dc5" = "Easier for disabled or older drivers",
      "benb_dc6" = "Saves money on drivers",
      "benb_dc7" = "Something else",
      "benb_dc8" = "None of these",
      "benb_dc9" = "Don't know"
    )
  ) +
  labs(
    title = "Perceived Benefits of Driverless Cars",
    x = NULL,
    y = "Percentage of Respondents (%)"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10),
    plot.title = element_text(face = "bold", size = 14)
  )
```

## Conclusion

-   statsiticaly significant

-   

## Next Steps
